### 2020-02-24 Draft1.tex
Initial commit of current pre-it progress

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{graphicx}

\setlength{\parindent}{4em}
\setlength{\parskip}{1em}

\title{Missing data and techniques for dealing with it}
\author{Marius Dioli}
\date{February 2019}

\begin{document}

	\maketitle

	\section{Introduction}
	If one is to do anything of import in science, one has to manipulate and analyze data. However, due to a variety of factors, data is often flawed. One major flaw one can have, and the topic of this thesis, is missing data. The consequence of missing data is always a reduction in efficiency and may also lead to bias.

	//A variety of ad hoc approaches are commonly used to deal with missing data, and many of the techniques described in this thesis are either//

	//This thesis will approach the problem of missing data from the bayesian perspective and with the goal of imputing or removing data. Other methods motivated from the frequentist perspective such as likelihood based methods will be touched on breifly as alternatives. The strength of the bayesian approach is that it provides a principled way to account for uncertainty about the missingness mechanism and the true distribution of the full data. //

	//Missing data is almost always considered in the context of fitting a model or making inference.//

	.. We will measure its efficacy on synthetic data and real data before comparing it with an alternate method that is gaining in popularity, machine learning.

	This paper is by no means comprehensive, but strives to give the reader a proper introduction to the problem of missing data and the technique of multiple imputation.
	// fyldig

	\section{Terminology and basic missingness theory}
	In the following section I will mostly adhere to the notation of Rubin in (citation), as it is both well established and quite general. However, I will adapt it slightly by using the observed/missing notation in (missing data) for readability, as it is easier for an unfamiliar audience and non-specialists to understand.

	\subsection{What is missingness?}
	When we say a variable is missing we mean that we have not observed it. The variable still has a value, but this value is hidden from us for some reason. What this reason might be will be described in detail later in this section. By treating missingness as a probabalistic phenomenon, statistically rigorous tools for dealing with it become available.

	Let $Y=(y_{ij})$ be a $nxk$ complete data matrix (i.e. with no missing values), with the ith row $y_{i}=(y_{i1}, ..., y_{ik})$ where $y_{ij}$ is the value of variable $Y_{j}$ for unit $i$. Let also the missingness indicator matrix $M = (m_{ij})$ be an $nxk$ such that $m_{ij} = 1$ if $y_{ij}$ is missing and $m_{ij} = 0$ if $y_{ij}$ is observed. Both $Y$ and $M$ are random variables, and we assume for simplicity that the rows $(y_{i}, m_{i})$ are independent and identically distributed over i. It is possible to have more than two possible values for each entry of the $M$, for example when we wish to differentiate between different kinds of non-response. However, we will not explore this in this paper. Now, given $m_{i}$, we can partition $y_{i}$ into $y^{o}_{i}$ and $y^{m}_{i}$, corresponding to the components that are observed and missing respectively (where the corresponding entry in $m_{i}$ is 0 or 1 respectively).

	\subsection{The missingness mechanism}

	The missingness mechanism is characterized by the conditional distribution of $m_{i}$ given $y_{i}$, i.e. $f_{M|Y}(m_{i}|y_{i}, \theta)$, where $\theta$ denotes unknown parameters. Stated differently the missingness mechanism is the probability of data being missing given the values of Y, missing and observed.

	The missingness pattern in $Y$ does not matter for the following definitions, and can be any of the following: *(Figure from page 9 Rubin)* 
	However, the missingness pattern will affect the efficacy of methods used to deal with missingness, and we shall explore this in subsequent sections.

	We distinguish between three different missingness mechanisms by considering how $M$ is related to $Y$: Missing Completely at Random (MCAR), Missing at Random (MAR), Missing Not at Random (MNAR). These can be roughly thought of as being in order of strictness from strictest (MCAR), to least strict (MNAR). Rubin(1976) [missing data book] introduced these missingness mechanisms, and they determine which techniques are appropriate to use on a given problem. **These mechanisms are rarely the case in practice, but it will be shown later that assuming them is reasonable in many circumstances. **

	\textbf{Missing Completely At Random (MCAR)}:
	This kind of missingness occurs when, $\forall i$ and any distinct values $y_{i}, y^{\ast}_{i}$ in the sample space of Y, $f_{M|Y}(m_{i}|y_{i}, \theta) = f_{M|Y}(m_{i}|y^{\ast}_{i}, \phi)$. In other words, it means that the probability of missingness is independent of both the observed and the missing data, and we can view the observed data as a random sample of the complete data.

	\textbf{Missing At Random (MAR)}:
	In this case, $\forall i$ and any distinct values $y^{m}_{i}, y^{m\ast}_{i}$ in the sample space of $y^{m}_{i}$, $f_{M|Y}(m_{i}|y^{o}_{i}, y^{m}_{i}, \theta) = f_{M|Y}(m_{i}|y^{o}_{i}, y^{m\ast}_{i}, \theta)$.
	This means that the probability of missingness, conditioned on the observed data, is independent of the missing data. Assuming MAR is less restrictive than MCAR, and may be considered a more plausible assumption about missing data in many contexts.

	\textbf{Missing Not At Random (MNAR)}:
	If the above equation does not hold for some $i$ and some $y^{m}_{i}, y^{m\ast}_{i}$, then the missingness mechanism is called missing not at random (MNAR). This means that the probability of missingness is dependent on the missing data even after conditioning on the observed data.

	\section{Illustrative example}

	The following example illustrates the three different types of missingness.

	Suppose you are doing a study in which you want to investigate people's sleep patterns. You send N people a survey asking them for the average amount of sleep they got each night in 2018. You send a follow-up survey for 2019. Table 1 shows simulated data for N = 15 people drawn from a bivariate normal distribution with $\mu_{x} = \mu_{y} = 7.5$, $\sigma_{x} = \sigma_{y} = 2$, and a correlation $\rho = 0.4$. The first two columns show the complete data for X and Y, while the other three columns show the values for 2019 after imposing missingness by three methods.

	\begin{figure}
		\centering
		\includegraphics[width=\linewidth]{missingness}
		\caption{Different missingness mechanisms}
		\label{fig:miss1}
	\end{figure}

	In the MCAR scenario, we see a random spread of missing data, i.e. a random subset of those who answered the 2018 survey also answered the 2019 one. Things get more interesting in the next two columns. Suppose you are only interested in those who've slept less than 7.5 hours on average, so you only send a follow-up survey to those who reported less than 7.5 hours of sleep for 2018 (2018<7.5). This is MAR, since the missingness of the 2019 entries depends on the values for 2018. For the final column, suppose you sent out your 2019 survey to all the individuals who answered the 2018 one, but include a line stating "Only answer this survey if you slept less than 7.5 hours on average in 2019" (2019 < 7.5). This is MNAR, since the missingness of the 2019 entries depends on the values for those entries. Remember, the full dataset exists, but it has been censored due to some mechanism, in this case through your own design.

	\section{The validity of missingness mechanisms}

	In controlled environemnts, like the one from the our example, we can guarantee that data are MAR since we control the censoring of the data (in this case by placing a cutoff between 2018 and 2019). However, it is rare that we have this much control over our data and a compelling reason to censor it so. In most real world scenarios we only have the data at hand to inform us about the missingnsess mechanism. Unfortunately, we cannot infer the missingness mechanism from data ...  We can check the validity of MCAR against MAR, but only under the unverifiable assumption that the data is not MNAR. Neither can we use the data at hand to support or refute on specific MNAR mechanism over another. The literature stresses the importance of sensitivity analyses in this case, checking the sensitivity of our inference under a variety of plausible assumptions. Finally MAR cannot be checked empirically against MNAR  (Cite from page 8, 9, and 10 in missing data). We therefore have to assume the missingness mechanism, most often MAR, and there are strong arguments for this. First and foremost is that assuming MAR yields good empirical results, which will be shown in later sections. %TODO Add more reasons for assuming MAR

	\section{What is imputation?}
	Imputation refers to substituting or ''filling in'' missing values with plausible estimates. The main strength of imputation is that it does not change the nature of the subsequent analysis. It makes it much easier for the data processor and analyst to work together since the processor's techniques won't interfere with the analysts'. Another benefit of imputation is that it augments the dataset, meaning that the dataset being analysed is larger than the original dataset, potentially yielding higher performance.

	\section{Techniques of interest}
	To properly evaluate the performance of MI, we should compare it to other similar and commonly used techniques, i.e. techniques that manipulate the dataset by either imputing or removing values. When choosing to apply a method it is important to keep two questions in mind. First, does the method under consideration give us consistent estimators for our model? Second, does the method give us appropriate measures of precision? How the following methods answer these questions is how we will decide the superiority of one method over another.

	\subsection{Ad hoc methods}
	Carpenter and Kenward (Missing data handbook) ...

	\textbf{Case deletion, also called complete case analysis (CCA):}
	A very common practice, CCA involves restricting analysis to entries with complete data. For those familiar with programming, a typical example would be to remove all rows from a dataframe with NaN entries. The first and most obvious problem with this approach is that it drastically reduces the amount of useable data. Additionally, this technique introduces large biases in the data if missingness is MAR or MNAR (CITE REVIEW PAPER FROM TURID). Only under MCAR is the data not biased when conducting CCA, but one is still faced with reduced accuracy due to missing data.

	\textbf{Mean imputation:}

	\textbf{Hot Deck imputation:}

	\subsection{Statistically principled methods}

	\textbf{Single imputation}

	Maximum likelihood and mixed linear models... This is the current alternative to the main method which we will discuss in the next chapter, multiple imputation. MI is often preferred over ML and MLM for two reasons: 
	1. It often performs similarly to multiple imputation with only a small advantage on certain kinds of data.
	2. It is less easy to use for non-statisticians
	A

	\section{Multiple imputation}

	Highly robust technique that yields gains even with relatively small portions of missing data.

	Only become feasible in the last 20 years due to computing power required.

	Bayesian approach.

	Between imputation variance

	\subsection{Specifying the imputation model}

	\subsection{Rubin's rules}

	\subsection{Potential pitfalls}
	When there is missing data in several variables of different types ,for example continuous and binary, and the data is structured, specifying the imputation model can be quite difficult. This is because it has to be a joint model...

	\subsection{When to use multiple imputation?}
	Computational resources and time to create the full joint imputation model. (Handbook 236)

	A common measure used to decide when to use multiple imputation is percentage of missing data.

	However, Madley-Dowd et. al. (Proportion of missing data paper) propose a different measure.

	MI can be used to perform sensitivity analysis by varying the imputation model.

	\section{Experiments}
	In this section we will use Madley-Dowd et. al.'s missingness measure to...

	\subsection{Synthetic example}

	"The MCAR missingness mechanism
	removed the first p observations such that pn gives the
	required proportion of missing data. MAR missingness
	was simulated under a logistic regression model using

	The value of a was manipulated for the different simu-
	lation settings to provide the required proportion of missing
	data on average across data sets." from (P. Madley-Dowd et a)

	Also check out their motivation for the generating of the data

	\subsubsection{Inference}
	Test statistics
	Different tests

	\subsubsection{Prediction}

	\subsubsection{Classification}

	\subsection{Performance across different data types}
	Continuous
	Categorical (binary, multicategory)
	Time to event
	Timeseries

	\subsection{Sensitivity analysis}

	\subsection{Real data}

	Description of the data

	HUNT Norwegian data

	Andreas data webSite
	balanced number of classes for classification

	\subsection{TODO}
	Difference between baysian and frequentist paradigm.
	Citation example:
	\cite[P.10]{grund_pooling_2016}.

	\bibliography{bib3}	
	\bibliographystyle{apalike}

\end{document}

### 2020-02-25 experiments.Rmd
Adding previous skeleton code

---
title: "Bachelor thesis"
author: "Marius Dioli"
date: "February 22, 2020"
output: pdf_document
---
```{r}

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(mice)
library(mvtnorm)
```

```{r}
#Setting vals for multivariate continous vars
mu = 8
sigma = 2
help(mvtnorm)
```

```{r}
#Generating binary data
n = 1000
binary_categorical_vector = round(runif(n, min = 0, max = 1))
```

### 2020-02-25 Draft1.tex
Minor changes

@@
 
-\title{Missing data and techniques for dealing with it}
+\title{Missing data: a simple introduction}
 \author{Marius Dioli}
@@
 
+	To quote MENG: ''imputation is not (merely) a compu-
+	tational tool but rather a mode of inference, which al-
+	lows hierarchical and sequential input of assessment
+	and information'' [Multiple-Imputation Inferences with Uncongenial Sources of Input]
+
 	//Missing data is almost always considered in the context of fitting a model or making inference.//
@@
 	// fyldig
+
+	As detailed in Xie and Meng [DISSECTING MULTIPLE IMPUTATION FROM], much of modern data analysis is done in a multi-phase paradigm, where the distinct and sequential phases data acquisition, data pre-processing, and data analysis, can have different or even contradictory assumptions.
 
@@
 
-	\section{Illustrative example}
+	\subsection{Illustrative example}
 
@@
 
-	\section{The validity of missingness mechanisms}
+	\subsection{The validity of missingness mechanisms}
 
@@
 
-	\section{What is imputation?}
+	\section{Imputation}
+
+	\subsection{What is imputation?}
 	Imputation refers to substituting or ''filling in'' missing values with plausible estimates.
@@
 It makes it much easier for the data processor and analyst to work together since the processor's techniques won't interfere with the analysts'.
+	%TODO change the previous sentence to reflect congeniality
 	 Another benefit of imputation is that it augments the dataset, meaning that the dataset being analysed is larger than the original dataset, potentially yielding higher performance.
 
-	\section{Techniques of interest}
+	\subsection{Congeniality: when the imputer's and analysts models differ}
+	As mentioned in the introduction, modern data analysis' multiple phases mean that the assumptions underlying each phase and the relationship between the phases is an important object of study.
+Meng's [meng 1994] concept of congeniality provides a theoretical framework from which to evaluate these differences.
+
+	Congeniality is an important concept in imputation and has a big impact on the statistical validity of MI.
+In short, uncongeniality arises when the analysis procedure differs substantially from the imputation model.
+There are many reasons for such difference, for example, the imputer might wish to include as many variables as possible to achieve an accurate imputation, while the analyst might be interested in only a subset of the covariates and the relationship between these.
+However, as Meng [1994] states, ''If the imputer's model is reasonably accu-
+	rate, then following the multiple-imputation recipe
+	prevents the analyst from producing inferences with
+	serious nonresponse biases''
+	Congeniality [Meng]
+
+	''When the imputation model class and the (embedded) analyst’s model class
+	differ, the behavior of Rubin’s rules becomes very complicated, capable of producing inconsistent variance estimators, a matter that has received recurrent
+	criticisms'' [Meng 2017]
+
+	\subsection{Techniques of interest}
 	To properly evaluate the performance of MI, we should compare it to other similar and commonly used techniques, i.e.
@@
 How the following methods answer these questions is how we will decide the superiority of one method over another.
+
+	There are two other competing methods which we should touch on breifly before continuing.
+Weighting procedures and likelihood-based methods are ...
+alternatives to multiple imputation.
+However, there is reason to prefer MI as the default options over these for a few reasons.
+First that, given congeniality, MI tends to outperform likelihood-based methods for small sample sizes, while delivering similar performance for large sample sizes \cite{schafer_multiple_2016}.
 
@@
 	Highly robust technique that yields gains even with relatively small portions of missing data.
-
-	Only become feasible in the last 20 years due to computing power required.
+It has only reached desired levels of performance in the last decade due to the computing power required.
 
 	Bayesian approach.
+Although MI is motivated from the Bayesian perspective, much of today's inference whether in science or industry is dominated by frequentist analyses.
+This causes some (Reference Fay and stuff from MENG) to question whether inference under MI is valid at all.
+To address this Meng [1994] developed the concept of congeniality, which we will go into more detail on soon.
 
-	Between imputation variance
+	Between imputation variance and within imputation variance
+
+	While MI is motivated from the bayesian perspective, it is most often used with subsequent frequentist inference. Thus ...
 
@@
 
-	\subsection{Rubin's rules}
+	We often want to include as many covariates as possible. X have shown that even mildly correlated auxiliary variables can improve subsequent inference [cite annotadet article]
 
-	\subsection{Potential pitfalls}
+	\subsection{Rubin's rules and Meng's adaptations}
+
+	Under congeniality, Rubin's rules for calculating standard errors, parameter values, and variance suffice.
+
+	If models are uncongenial, we use Meng's rules.
+
+	\subsection{Potential pitfalls and criticism}
 	When there is missing data in several variables of different types ,for example continuous and binary, and the data is structured, specifying the imputation model can be quite difficult.
 This is because it has to be a joint model...
+
+	. Among all the
+	criticisms of multiple imputation, the recent work
+	of Fay (1991, 1992) is the most intense, as it is
+	directed at the validity of multiple-imputation in-
+	ference in practice. Fay provided several examples
+	(see also Kott, 1992) to show that the variance esti-
+	mator obtained from the repeated-imputation combining rules disagrees asymptotically with the sam-
+	pling variance of the repeated-imputation estimator,
+	even when the imputation model is correctly speci-
+	fied. [MENG]
+
+	%^fixed with congeniality
+
+	Taking all this into account, it is important to see MI in context of the options available to someone working with data.
+As we have established earlier, simply ignoring missing data can lead  to serious biases, while alternative methods yield similar if not inferior performance in certain cases.
+As long as a potential user is aware of the potential pitfalls, MI is the superior method for dealing with missing data.
 
@@
 
+	\subsection{MICE}
+	''The
+	increased flexibility in modeling these conditional distributions may outweigh
+	the lack of clear theoretical justification of the method.'' Rubin
+
 	\subsection{Synthetic example}
@@
 
-	Also check out their motivation for the generating of the data
+	Also check out their motivation for the generating of the dataa
+
+	In the following examples we will compare our models' performance across different data types.
+These are continuous, categorical (binary and multicategory), time to event data, and timeseries.
+The latter two
 
@@
 	Different tests
+	x
+	Congenial vs uncongenial models, see the annotated paper for how they did things with regard to imputation vs analysis.
 
 	\subsubsection{Prediction}
+	With large publicly available datasets on the one hand, and laws like GDPR requiring deleting certain kinds of data on the other, multiple imputation seems like a strong candidate for improving the predictive quality of models.
+Most of the literature focuses on inference, so there is comparatively little comparing the performance of different predictive models.
+Here we will compare three basic model architectures, multiple linear regression, boosted trees, and a simple neural network, to see how they perform with and without multiple imputation.
 
 	\subsubsection{Classification}
+	Like with the prediction case above, classification is an area in which the litterature on missing data is sparse and the same conditions apply.
+In this context we will compare logistic regression, boosted trees, and neural networks.
 
-	\subsection{Performance across different data types}
-	Continuous
-	Categorical (binary, multicategory)
-	Time to event
-	Timeseries
-
-	\subsection{Sensitivity analysis}
+	\subsection{Sensitivity analysis and uncongenial models}
 
@@
 
-	\bibliography{bib3}	
+	\bibliography{bib2}	
 	\bibliographystyle{apalike}

### 2020-02-26 Draft1.tex
Many small additions. Theory is coming together

@@
 	\section{Introduction}
+	... ''eng and Romero (2003), Xie and Meng (2012)). In general,
+	the problem of multiphase inference can be formulated as one of missing data. However, in the
+	multiphase setting, missingness arises from the preprocessing choices made, not a probabilis
+	tic response mechanism. Thus, we can leverage the mathematical and computational methods
+	of this literature, but many of its conceptual tools need to be modified. Multiple imputation ad
+	dresses many of the same issues as multiphase inference and is indeed a special case of the
+	latter. Concepts such as congeniality between imputation and analysis models and self-efficiency
+	(Meng ( 1994)) have natural analogues and roles to play in the analysis of multiphase inference
+	problems.''
+
+	In this paper we will focus on the Multiple Imputation in the context of correcting for missing data before analysis.
+Blocker and Meng argue that general preprosessing can be evaluated in the missing data paradigm, but this is outside the scope of this paper.
+We refer readers to Block and Meng for an in depth consideration of the subject.
+
+	The public database is perhaps the context where MI is most useful.
+These are often analyzed by multiple users with varying degrees of statistical knowledge and resources.
+All public databases have missing data, and typically users do not have the knowledge to deal with missing data satisfactoraly.
+Even if they do, the database constructors often have more information about the missing data than the end user, and so can better compensate for this.
+	%TODO Introduce the concept of multiphase inference. Reference meng2017 and Blocker and meng
 	If one is to do anything of import in science, one has to manipulate and analyze data.
@@
 
-	In controlled environemnts, like the one from the our example, we can guarantee that data are MAR since we control the censoring of the data (in this case by placing a cutoff between 2018 and 2019).
-However, it is rare that we have this much control over our data and a compelling reason to censor it so.
-In most real world scenarios we only have the data at hand to inform us about the missingnsess mechanism.
-Unfortunately, we cannot infer the missingness mechanism from data ...
+	In controlled environments, like the one from our example, we can guarantee that data are MAR since we control the censoring of the data (in this case by placing a cutoff between 2018 and 2019).
+However, it is rare that we have this much control over our data, and a compelling reason to censor it so.
+In most real world scenarios we only have the data at hand to inform us about the missingness mechanism.
+Unfortunately, we cannot infer the missingness mechanism from data.
 We can check the validity of MCAR against MAR, but only under the unverifiable assumption that the data is not MNAR.
@@
 	\subsection{What is imputation?}
-	Imputation refers to substituting or ''filling in'' missing values with plausible estimates.
-The main strength of imputation is that it does not change the nature of the subsequent analysis.
+	Imputation refers to substituting or ''filling in'' of missing values with plausible estimates.The main strength of imputation is that it does not change the nature of the subsequent analysis.
 It makes it much easier for the data processor and analyst to work together since the processor's techniques won't interfere with the analysts'.
-	%TODO change the previous sentence to reflect congeniality
-	 Another benefit of imputation is that it augments the dataset, meaning that the dataset being analysed is larger than the original dataset, potentially yielding higher performance.
+This is, at least, the goal of imputation.
+In certain situation the information available to the imputer might differ to that available to the analyst.
+In such cases difficulties arise.
+The concept of congeniality, which we will define shortly, helps us overcome some of these difficulties.
+Another very practical benefit, and a strong argument for its use, is the performance gains to be had from using it.
+This will become clear in the experiments section.
 
@@
 Meng's [meng 1994] concept of congeniality provides a theoretical framework from which to evaluate these differences.
+
+	''When the imputation model class and the (embedded) analyst’s model class
+	differ, the behavior of Rubin’s rules becomes very complicated, capable of producing inconsistent variance estimators, a matter that has received recurrent
+	criticisms'' [Meng 2017]
 
@@
 	serious nonresponse biases''
-	Congeniality [Meng]
 
-	''When the imputation model class and the (embedded) analyst’s model class
-	differ, the behavior of Rubin’s rules becomes very complicated, capable of producing inconsistent variance estimators, a matter that has received recurrent
-	criticisms'' [Meng 2017]
+	\subsection{Does imputation yield valid inference?}
+	Techniques are baased on the bayesian paradigm, but
+
+	Rubin (1996) defines
 
@@
 
-	There are two other competing methods which we should touch on breifly before continuing.
+	There are two other competing methods which we should touch on briefly before continuing.
 Weighting procedures and likelihood-based methods are ...
@@
 	\textbf{Case deletion, also called complete case analysis (CCA):}
-	A very common practice, CCA involves restricting analysis to entries with complete data.
+	Probably the most common technique in use for dealing with missing data, CCA involves restricting analysis to entries with complete data.
 For those familiar with programming, a typical example would be to remove all rows from a dataframe with NaN entries.
 The first and most obvious problem with this approach is that it drastically reduces the amount of useable data.
-Additionally, this technique introduces large biases in the data if missingness is MAR or MNAR (CITE REVIEW PAPER FROM TURID).
-Only under MCAR is the data not biased when conducting CCA, but one is still faced with reduced accuracy due to missing data.
+This holds an intuitive appeal, as it is both exceedingly simple to implement, and seem to have the regrettable but sometimes minor effect of lower accuracy.
+However, depending on the missingness mechanism, this might be wrong, and can actually introduce bias into the data.
+For example, suppose you want to find the mean and standard deviation in the height of a population.
+If for some reason you lack all entries for people above 200cm, then disregarding these missing entries in a CCA will lead to biased estimates for both the mean and variance.
+An implicit assumption, therefore, of CCA is that the observed data is a random sample of the underlying distribution.
+This means that CCA implicitly assumes MCAR, and we see in [schafer and graham] that we get unbiased inference in this scenario, if with lower accuracy depending on the amount of missing data.
+However, as with the height example, deviation from MCAR can introduce bias.
+(CITE REVIEW PAPER FROM TURID).
 
 	\textbf{Mean imputation:}
+	Another common practice, this technique involves substituting the missing values with the mean for that variable.
+As stated in Schafer and Graham [2002], this has the effect of preserving the mean of the distribution one wishes to draw inference about, but distorts other parameters such as variance and covariance.
+%Add some more math from either schafer and graham or another source
 
@@
 
+	\textbf{Last Observation Carried Forward (LOCF):x}
+
 	\subsection{Statistically principled methods}
+	The bayesian paradigm is ideal for more statistically principled imputation, as it allows us to effectively model the unknown given the known, and makes assumptions explicit.
 
@@
 
-	Maximum likelihood and mixed linear models...
-This is the current alternative to the main method which we will discuss in the next chapter, multiple imputation.
-MI is often preferred over ML and MLM for two reasons:
-	1. It often performs similarly to multiple imputation with only a small advantage on certain kinds of data.
-	2. It is less easy to use for non-statisticians
-	A
+	\section{Multiple imputation}
 
-	\section{Multiple imputation}
+	Multiple imputation can be thought of as three distinct steps:
+	1) M imputed datasets are generated
+	2) The intended analysis is performed separately on all M datasets
+	3) The results of the analysis are combined using Rubin's or Meng's rules to create a single MI result.
 
@@
 
-	We often want to include as many covariates as possible. X have shown that even mildly correlated auxiliary variables can improve subsequent inference [cite annotadet article]
+	We often want to include as many covariates as possible.
+X have shown that even mildly correlated auxiliary variables can improve subsequent inference [cite annotadet article].
+However, this can cause the imputation and and analyses models to be uncongenial.
 
@@
 	\subsection{Potential pitfalls and criticism}
+	The following arguments apply for imputation in general, but we will consider them in the context of MI.
 When there is missing data in several variables of different types ,for example continuous and binary, and the data is structured, specifying the imputation model can be quite difficult.
@@
 	the lack of clear theoretical justification of the method.'' Rubin
+
+	We have chosen to use MICE for the experiments in this paper for two main reasons.
+One is the aformentioned performance and flexibility, while the other is that MICE is unlicenced and open source in contrast to other technqieus like drawing from a fully specified joint distribution found in statistical packages like STATA.
+In the interest of democratizing MI while retaining the desired performance, MICE was found to be superior.
 

### 2020-02-27 Draft1.tex
adding a ref on mice
Finished ad hoc methods

@@
 	\subsection{Ad hoc methods}
-	Carpenter and Kenward (Missing data handbook) ...
+	This section is mostly adapted from Carpenter and Kenward [Ch2.5 missing data].
+Although Ad hoc is pejorative in certain circumstances, it is used here to indicate that the methods below are motivated by convenience rather than any methodological concern.
 
@@
 However, as with the height example, deviation from MCAR can introduce bias.
+See Schafer and Graham, ...
+for examples of how inference is affected.
 (CITE REVIEW PAPER FROM TURID).
@@
 	\textbf{Mean imputation:}
-	Another common practice, this technique involves substituting the missing values with the mean for that variable.
+	Another common practice, this technique involves substituting the missing values with the mean of the observed values for that variable.
 As stated in Schafer and Graham [2002], this has the effect of preserving the mean of the distribution one wishes to draw inference about, but distorts other parameters such as variance and covariance.
+Covariance especially is distorted due to mean imputation taking only the varible under consideration as input.
+In addition to this the technique breaks down when confronted with categorical variables where the mean does not exist (and it is clear how a majority vote imputation might skew the data).
 %Add some more math from either schafer and graham or another source
 
-	\textbf{Hot Deck imputation:}
+	\textbf{Last Observation Carried Forward (LOCF):x}
+	This method is mostly used in longitudinal studies where individuals drop out, withdraw, or are lost over the course of the studies.
+LOCF involves imputing the missing values for an individual with the last observed value for that individual.
+Leaving aside questions of precision, LOCF produces biased estimators in all except very specific MNAR circumstances (Kenward and Molenberghs 2009 get article).
+With LOCF we are effectively assuming that for each individual that drops out their responses do not change at all for the remainder of the study, a highly unrealistic assumption.
+It can be concluded, therefore, that LOCF is the worst imputation method of those mentioned here, and all the other methods are preferable.
+According [Missing data] "LOCV is neither valid under general assumptions nor based on statistical principles, it is not a sensible method, and should not be used".
 
-	\textbf{Last Observation Carried Forward (LOCF):x}
+	\textbf{Regression:}
+	Regression imputation functions by regressing the variable(s) with missing entries on the fully observed variables, or subsets that are fully observed.
+To illustrate this, let us consider the simplest case, with one fully observed variable $x$ linearly related to a variable with missing data $y$.
+Fitting a regression of $y$ on $x$, we have$$Y_{i} = \beta_{0} + \beta_{1}x_{i} + \epsilon \qquad i\in(0,...,n)$$ obtaining estimates of the regression parameters $\hat{\beta_{0}}$ and $\hat{\beta_{1}}$ allows us to impute $E(y_{i}) = \hat{\beta_{0}} + \hat{\beta_{1}}x_{i}$.
+This is the best of the ad hoc methods, as it gives us much more plausible imputations that can vary.
+However, even in this simple case it yields imputations that tend to be less variable than the observed data.
+Therefore, using them in inference will most often lead to underestimated standard errors and p-values.
+If there are non-linear relationships in the data, this method's performance breaks down with regard to other aspects of inference as well.
+
+	For more in depth descriptions and examples of these methods we refer the reader to [missing data handbook].
 
@@
 	\textbf{Single imputation}
+	To build our intuition about multiple imputation it is relevant to discuss the single imputation case, and how its shortcomings are rectified in MI.
 
 	\section{Multiple imputation}
+	One sizeable weakness of the earlier imputation methods is that they use points as estimates of distributions.
+Hence, subsequent measures of precision are biased downwards.
+Multiple imputation solves this problem by creating multiple imputed datasets drawn from a posterior distribution and then combines them in the final step.
+This preserves the distributional variation in the missing data, and
 

### 2020-02-27 Draft1.tex
Minor additions to statistical validity

@@
 
-	//Missing data is almost always considered in the context of fitting a model or making inference.//
-
 	.. We will measure its efficacy on synthetic data and real data before comparing it with an alternate method that is gaining in popularity, machine learning.
@@
 	As detailed in Xie and Meng [DISSECTING MULTIPLE IMPUTATION FROM], much of modern data analysis is done in a multi-phase paradigm, where the distinct and sequential phases data acquisition, data pre-processing, and data analysis, can have different or even contradictory assumptions.
+
+	Our objective with multiple imputation, as stated in Rubin(1996), is: assuming that ultimate user's complete-data analysis is stat
+
+	Although our main objective with MI is inference, we will test its ability to improve performance in prediction and classification contexts as well.
+
+	The goal of this paper is to give an overview of missingness theory and multiple imputation such that as many as possible can then apply MI well.
 
@@
 	\subsection{Does imputation yield valid inference?}
-	Techniques are baased on the bayesian paradigm, but
+	Techniques are based on the bayesian paradigm, but
 
-	Rubin (1996) defines
+	Rubin (1996) defines the term scientific estimand as the objects we wish to draw inference about. He defines them thusly: "... a quantity of scientific
+	interest that can be calculated in the population and does
+	not change its value depending on the data collection de-
+	sign used to measure it (i.e., it does not vary with sample
+	size and survey design, or the number of nonrespondents,
+	or follow-up effort. Sci-
+	entific estimands include population means, variances, cor-
+	relations, factor loadings, regression coefficients, and these
+	quantities within strata or domains, but exclude the sam-
+	pling variance of a sample mean under a particular sampling
+	plan and the expectation of the complete-data sample mean
+	when missing values are filled in with zero or the observed sample mean.
+".
+Rubin goes on to describe statistical validity for scientific estimands as the following: " point estimation must be approximately unbi-
+	ased for the scientific estimands, averaging over the sam-
+	pling and the posited nonresponse mechanisms (e.g., filling
+	in zeros or means is not generally acceptable)." and "in-
+	terval estimation and hypothesis testing must be valid in
+	the sense that nominal levels describe operating character-
+	istics over sampling and posited response mechanisms.".
+The latter criteria can be further refined into "randomization validity" and "confidence validity".
+The first is that, according to Rubin (1996),  "randomization validity means that, for interval estimates,
+	"actual interval coverage = nominal interval coverage," and
+	for tests of hypotheses, "actual rejection rate = nominal re-
+	jection rate."  Randomization validity is of course desireable, but not always achievable, and so Rubin introduces confidence validity as a looser criterion and defines it as "hat for interval estimates, "ac-
+	tual interval coverage > nominal interval coverage," and for
+	tests of hypotheses, "actual rejection rate < nominal rejec-
+	tion rate.""
 
@@
 
+	Lee and Carlin  [2010] argue that MICE and MVN have comparable performance.
+
 	We have chosen to use MICE for the experiments in this paper for two main reasons.

### 2020-02-28 Draft1.tex
Minor changes

@@
 
-	\subsection{Congeniality: when the imputer's and analysts models differ}
-	As mentioned in the introduction, modern data analysis' multiple phases mean that the assumptions underlying each phase and the relationship between the phases is an important object of study.
-Meng's [meng 1994] concept of congeniality provides a theoretical framework from which to evaluate these differences.
+	The more advanced model based approaches of this paper uses both the Bayesian and frequentist paradigms in complementary ways.
+The Bayesian approach creates procedures, while the frequentist one evalueates them.
 
-	''When the imputation model class and the (embedded) analyst’s model class
-	differ, the behavior of Rubin’s rules becomes very complicated, capable of producing inconsistent variance estimators, a matter that has received recurrent
-	criticisms'' [Meng 2017]
-
-	Congeniality is an important concept in imputation and has a big impact on the statistical validity of MI.
-In short, uncongeniality arises when the analysis procedure differs substantially from the imputation model.
-There are many reasons for such difference, for example, the imputer might wish to include as many variables as possible to achieve an accurate imputation, while the analyst might be interested in only a subset of the covariates and the relationship between these.
-However, as Meng [1994] states, ''If the imputer's model is reasonably accu-
-	rate, then following the multiple-imputation recipe
-	prevents the analyst from producing inferences with
-	serious nonresponse biases''
-
-	\subsection{Does imputation yield valid inference?}
-	Techniques are based on the bayesian paradigm, but
+	\subsection{What is valid inference?}
+	Techniques are based on the bayesian paradigm, but inference is often done in the frequentist one.
 
@@
 	jection rate."  Randomization validity is of course desireable, but not always achievable, and so Rubin introduces confidence validity as a looser criterion and defines it as "hat for interval estimates, "ac-
-	tual interval coverage > nominal interval coverage," and for
-	tests of hypotheses, "actual rejection rate < nominal rejec-
-	tion rate.""
+	tual interval coverage >= nominal interval coverage," and for
+	tests of hypotheses, "actual rejection rate =< nominal rejec-
+	tion rate."" We therefore proceed in our analysis of methods using these definitions when evaluating the inferential validity of imputation methods.
+
+	\subsection{Congeniality: when the imputer's and analysts models differ}
+	As mentioned in the introduction, modern data analysis' multiple phases mean that the assumptions underlying each phase and the relationship between the phases is an important object of study.
+Meng's [meng 1994] concept of congeniality provides a theoretical framework from which to evaluate these differences.
+
+	''When the imputation model class and the (embedded) analyst’s model class
+	differ, the behavior of Rubin’s rules becomes very complicated, capable of producing inconsistent variance estimators, a matter that has received recurrent
+	criticisms'' [Meng 2017]
+
+	Congeniality is an important concept in imputation and has a big impact on the statistical validity of MI.
+In short, uncongeniality arises when the analysis procedure differs substantially from the imputation model.
+There are many reasons for such difference, for example, the imputer might wish to include as many variables as possible to achieve an accurate imputation, while the analyst might be interested in only a subset of the covariates and the relationship between these.
+However, as Meng [1994] states, ''If the imputer's model is reasonably accu-
+	rate, then following the multiple-imputation recipe
+	prevents the analyst from producing inferences with
+	serious nonresponse biases''
 
@@
 
+	Even Rubin (1996) admits that MI
+
 	Taking all this into account, it is important to see MI in context of the options available to someone working with data.
@@
 	\subsection{Synthetic example}
+
+	%Remember to use Rubin's definitons of statistical validity in evaluating the methods.
 

### 2020-03-02 Draft1.tex
Some changes to congeniality

@@
 
-	Let $Y=(y_{ij})$ be a $nxk$ complete data matrix (i.e.
+	Let $Y=(y_{ij})$ be a $n*k$ complete data matrix (i.e.
 with no missing values), with the ith row $y_{i}=(y_{i1}, ..., y_{ik})$ where $y_{ij}$ is the value of variable $Y_{j}$ for unit $i$.
@@
 The Bayesian approach creates procedures, while the frequentist one evalueates them.
+This is because the bayesian paradigm allows us to effectively model the unknown given the known, and makes assumptions explicit.
+The frequentist paradigm is the one which is by far most common when doing inference, so it makes sense that it is under that paradigm performance matters.
 
@@
 
+	MI is motivated under the Bayesian paradigm.
+This causes some (Reference Fay and stuff from MENG) to question whether inference under MI is valid at all.
+To address this Meng [1994] developed the concept of congeniality, which we will go into more detail on soon.
+
 	\subsection{Congeniality: when the imputer's and analysts models differ}
@@
 Meng's [meng 1994] concept of congeniality provides a theoretical framework from which to evaluate these differences.
+
+	First let us define some of the basic notation that Meng (1994) uses to define congeniality.
+We will modify it slightly to fit with the notation of this paper and since Meng's paper was written in the context of surveys.
+An analyst's complete-data procedure can be summarised by $P_{com} = {\hat{Q}(Y, M), U(Y, M)}$ where $\hat{Q}(Y, M)$ is an estimator of $Q(Y, M)$, an uknown quantity of interest, with $U(Y, M)$ being an associated variance estimator.
+Likewise, an analyst's incomplete data procedure is denoted by $P_{obs} = {\hat{Q}(Y^{o}, M), U(Y{o}, M)}$.
+Meng goes on to define congeniality in the following manner: "A Bayesian model f is said to be congenial to the analysis procedure $P = {P_{obs};P_{com}}$ for a given ${Y^{o}, M}$ if the following hold:
+
+	1) The posterior mean and variance of $Q$ under
+	f given the incomplete data are asymptotically the
+	same as the estimate and variance from the analyst's
+	incomplete-data procedure , $P_{obs}$ that is, $$[[\hat{Q}(Y^{o}, M), U(Y^{o}, M)]\simeq [E_{f}[Q|{Y^{o}, M}], Var_{f}[Q|{Y^{o}, M}]]$$
+
+	2)  The posterior mean and variance of $Q$ under
+	f given the complete data are asymptotically the
+	same as the estimate and variance from the analyst's
+	complete-data procedure , $P_{com}$ that is, $$[[\hat{Q}(Y), U(Y)]\simeq [E_{f}[Q|{Y}], Var_{f}[Q|{Y}]]\qquad \forall Y\textrm{with}Y^{o}\textrm{fixed (i.e., conditioned upon)}$$
 
@@
 
-	\subsection{Statistically principled methods}
-	The bayesian paradigm is ideal for more statistically principled imputation, as it allows us to effectively model the unknown given the known, and makes assumptions explicit.
+	\subsection{Statistically principled imputation}
+	By statistically principled, we mean motivated by an ability to preserves the statistical qualities in the data which we care about.
+For inference, this would be parameter values of a distribution and their variance and covariance.
 
 	\textbf{Single imputation}
-	To build our intuition about multiple imputation it is relevant to discuss the single imputation case, and how its shortcomings are rectified in MI.
+	Single imputation is the building block of MI, and ..
+Single imputation can be done in a variety of ways, but the most straightforward and easy to understand is the following: create a joint posterior distribution ...
+Draw from this distribution to fill in your dataset, and then do your analysis.
 
@@
 Multiple imputation solves this problem by creating multiple imputed datasets drawn from a posterior distribution and then combines them in the final step.
-This preserves the distributional variation in the missing data, and
+This preserves the distributional variation in the missing data, and ...
+Increases in available computation power has made MI a much more practical method to implement, as one of the weaknesses that earlier impeded its adoption was that it was computationally expensive.
 
 	Multiple imputation can be thought of as three distinct steps:
-	1) M imputed datasets are generated
+	1) M imputed datasets are generated through repeated random draws from the predictive distribution of the missing values under a particular model for missingness.
 	2) The intended analysis is performed separately on all M datasets
@@
 
-	Highly robust technique that yields gains even with relatively small portions of missing data.
-It has only reached desired levels of performance in the last decade due to the computing power required.
-
-	Bayesian approach.
-Although MI is motivated from the Bayesian perspective, much of today's inference whether in science or industry is dominated by frequentist analyses.
-This causes some (Reference Fay and stuff from MENG) to question whether inference under MI is valid at all.
-To address this Meng [1994] developed the concept of congeniality, which we will go into more detail on soon.
-
 	Between imputation variance and within imputation variance
 
-	While MI is motivated from the bayesian perspective, it is most often used with subsequent frequentist inference. Thus ...
+	It should be noted that even though there is no strict mathematical proof that MI yields valid inference, the empirical evidence presents a strong case that MI is a useful tool.
+Comparing it to the other methods available, one finds that alternatives are on the whole less effective.
 

### 2020-03-03 Draft1.tex
Finished defining congeniality and added Rubin's rules

@@
 	As mentioned in the introduction, modern data analysis' multiple phases mean that the assumptions underlying each phase and the relationship between the phases is an important object of study.
-Meng's [meng 1994] concept of congeniality provides a theoretical framework from which to evaluate these differences.
+Meng's [meng 1994] concept of congeniality provides a theoretical framework from which to evaluate these differences and most of this section is derived from his paper.
 
@@
 	complete-data procedure , $P_{com}$ that is, $$[[\hat{Q}(Y), U(Y)]\simeq [E_{f}[Q|{Y}], Var_{f}[Q|{Y}]]\qquad \forall Y\textrm{with}Y^{o}\textrm{fixed (i.e., conditioned upon)}$$
+
+	Additionally, let $g(Y^{m}|{Y^{o}, M}, A)$ be the imputation model, where A represents possible additional data the imputer has access to.
+Meng continues by stating ''The analysis procedure $P$ is said to be congenial to the imputation model $g(Y^{m}|{Y^{o}, M}, A)$ if one can find an f such that 1) f is congenial to $P$ and 2) the posterior predictive density for $Y^{m}$ derivd under f is identical to the imputation model $$f(Y{m}|{Y^{o}, M}) = g(Y{m}|{Y^{o}, M}, A) \forall Y^{m}$$
+
+	Therefore if no such congenial f exists, then$P$ is said to be uncongenial to g.
+Notice that congeniality is defined with respect to a particular analysis procedure, meaning that an analyst can perform two types of analyses where one is congenial and the other is not.
+There are three main scenarios in which uncongeniality occurs in the context of missing data and MI.
+The first is when the imputation model is largely unknown to the analyst, often the case with public databases.
+Second is when the different needs of imputation and analysis require different models.
+Third is when one is conducting sensitivity analysis, which necessarily requires the testing of different models, some of which will be uncongenial.
 
@@
 
-	Under congeniality, Rubin's rules for calculating standard errors, parameter values, and variance suffice.
+	At step 2) of MI, we have M sets of parameter estimates and their variances $\hat{\theta}_{i},\hat{Var_{i}} \qquad i=0,...,M$.
+The next step is to combine these into a single estimate.
+Under congeniality Rubin's combining rules suffice for giving us statistically valid estimates.
+The parameter estimate is simply the average $$\bar{\theta} = \frac{1}{M}\sum_{i=0}^{M}\hat{\theta}_{i}$$ and the variance associated with $\bar{\theta}$ is given by $$T = \bar{Var}+(1+\frac{1}{M}B)$$ where $$\bar{Var} = \frac{1}{M}\sum_{i=0}^{M}\hat{Var}_{i}$$ measures the within imputation variability and $$B = \frac{1}{M-1}\sum_{i=0}^{M}(\hat{\theta_{i}}-\bar{\theta})(\hat{\theta_{i}}-\bar{\theta})^{T}$$ measures the between imputation variability.
+The adjustment $(1+\frac{1}{M})$ is due to the finite number of imputations.
+If we let $M\rightarrow\infty$ then...
 

### 2020-03-03 Draft1.tex
Finished a first draft of all major theory sections

@@
 	\subsection{What is imputation?}
-	Imputation refers to substituting or ''filling in'' of missing values with plausible estimates.The main strength of imputation is that it does not change the nature of the subsequent analysis.
+	Imputation refers to substituting or ''filling in'' of missing values with plausible estimates.
+Often the person creating or maintaining a dataset will have access to more information than the one analysing it.
+Using this information in correcting for missing data means that the imputer can do a better job than the analyst.
+The main strength of imputation is that it does not change the nature of the subsequent analysis.
 It makes it much easier for the data processor and analyst to work together since the processor's techniques won't interfere with the analysts'.
@@
 	\subsection{What is valid inference?}
-	Techniques are based on the bayesian paradigm, but inference is often done in the frequentist one.
+	Some have questioned whether methods developed under the Bayesian paradigm leads to valid frequentist inference (Meng1994).
+However, Xia and Meng(2017), Meng(1994), and Rubin(1996) have shown that such Bayesian models yield valid inference in the asymptotic case, and even in the non-asymptotic case following the below definition given certain conditions which we will explore later.
 
@@
 
-	MI is motivated under the Bayesian paradigm.
-This causes some (Reference Fay and stuff from MENG) to question whether inference under MI is valid at all.
-To address this Meng [1994] developed the concept of congeniality, which we will go into more detail on soon.
-
 	\subsection{Congeniality: when the imputer's and analysts models differ}
 	As mentioned in the introduction, modern data analysis' multiple phases mean that the assumptions underlying each phase and the relationship between the phases is an important object of study.
+Differences might cause the subsequent analysis to be biased, or in the worst case, invalid.
 Meng's [meng 1994] concept of congeniality provides a theoretical framework from which to evaluate these differences and most of this section is derived from his paper.
@@
 	Additionally, let $g(Y^{m}|{Y^{o}, M}, A)$ be the imputation model, where A represents possible additional data the imputer has access to.
-Meng continues by stating ''The analysis procedure $P$ is said to be congenial to the imputation model $g(Y^{m}|{Y^{o}, M}, A)$ if one can find an f such that 1) f is congenial to $P$ and 2) the posterior predictive density for $Y^{m}$ derivd under f is identical to the imputation model $$f(Y{m}|{Y^{o}, M}) = g(Y{m}|{Y^{o}, M}, A) \forall Y^{m}$$
+Meng continues by stating ''The analysis procedure $P$ is said to be congenial to the imputation model $g(Y^{m}|{Y^{o}, M}, A)$ if one can find an f such that 1) f is congenial to $P$ and 2) the posterior predictive density for $Y^{m}$ derived under f is identical to the imputation model $$f(Y{m}|{Y^{o}, M}) = g(Y{m}|{Y^{o}, M}, A) \forall Y^{m}$$
 
@@
 
-	''When the imputation model class and the (embedded) analyst’s model class
-	differ, the behavior of Rubin’s rules becomes very complicated, capable of producing inconsistent variance estimators, a matter that has received recurrent
-	criticisms'' [Meng 2017]
-
 	Congeniality is an important concept in imputation and has a big impact on the statistical validity of MI.
@@
 	prevents the analyst from producing inferences with
-	serious nonresponse biases''
+	serious nonresponse biases''.
+Meng goes on to show that even with an uncongenial imputation model, inferences performed on a MI dataset still yield better inference than the best possible incomplete data inference.
 
@@
 	\textbf{Single imputation}
-	Single imputation is the building block of MI, and ..
-Single imputation can be done in a variety of ways, but the most straightforward and easy to understand is the following: create a joint posterior distribution ...
-Draw from this distribution to fill in your dataset, and then do your analysis.
+	Single imputation is the building block of MI, and many of it's deficiencies are corrected for when expanded to MI.
+It can be done in a variety of ways, but the most straightforward and easy to understand is the following: create a joint posterior distribution of your missing data given your observed data, i.e.
+$f(Y^{m}|Y^{o}, M)$.
+Draw from this distribution to fill in your dataset, then perform your desired analysis.
+How you specify the missing data model can determine the efficacy of this method, but it is otherwise uncomplicated.
+We will therefore focus the majority of our analysis on the MI case.
 
@@
 Multiple imputation solves this problem by creating multiple imputed datasets drawn from a posterior distribution and then combines them in the final step.
-This preserves the distributional variation in the missing data, and ...
+This preserves the distributional variation in the missing data, and leads to better variance estimates.
 Increases in available computation power has made MI a much more practical method to implement, as one of the weaknesses that earlier impeded its adoption was that it was computationally expensive.
+Many of the reservations people had with MI regarding inference concerned MI with few imputed datasets M, typically in the 5-10 range.
+With larger M and more advanced theoretical framework for combining the resulting analyses, one can achieve strong results in a large number of situations, see Xia and Meng (2017), Meng (1994), Rubin (1996), and Madley-Dowd et al.
+(2019).
 
 	Multiple imputation can be thought of as three distinct steps:
-	1) M imputed datasets are generated through repeated random draws from the predictive distribution of the missing values under a particular model for missingness.
+	1) M imputed datasets are generated through repeated random draws from the predictive distribution of the missing values ($f(Y^{m}|Y^{o}, M)$) under a particular model for missingness.
 	2) The intended analysis is performed separately on all M datasets
@@
 
-	Between imputation variance and within imputation variance
+	It should be noted that even though there is no strict mathematical proof that MI yields valid inference in the case of finite M, the empirical evidence presents a strong case that MI is a useful tool.
+Comparing it to the other methods available, one finds that alternatives are on the whole less effective.
+%Additionally, Meng (1994) shows us that for sufficiently many imputations we have that the variance is less than for one calculated in the complete case.
 
-	It should be noted that even though there is no strict mathematical proof that MI yields valid inference, the empirical evidence presents a strong case that MI is a useful tool.
-Comparing it to the other methods available, one finds that alternatives are on the whole less effective.
+	%If MI is used, the imputer should convey as clearly as possible to the analyst the form of their model and its underlying assumptions.
 
@@
 
-	We often want to include as many covariates as possible.
-X have shown that even mildly correlated auxiliary variables can improve subsequent inference [cite annotadet article].
-However, this can cause the imputation and and analyses models to be uncongenial.
+	To create an accurate model for the missing data, we often want to include as many covariates as possible.
+Madley-Dowd et al.
+(2019) have shown that even mildly correlated auxiliary variables can improve subsequent inference.
+This leads us to make repeated random draws from $f(Y^{m}|Y^{o}, M, A)$, where A represents auxiliary variables.
+Whether or not to include auxialiary variables is a judgement best made by the imputer, as it can both be a cause and a cure for uncongeniality.
+Meng (1994) cautions against the use of less general imputation models (e.g.
+ignoring covariates/auxiliary variables).
 
-	\subsection{Rubin's rules and Meng's adaptations}
+	\subsection{Rubin's rules}
 
@@
 The adjustment $(1+\frac{1}{M})$ is due to the finite number of imputations.
-If we let $M\rightarrow\infty$ then...
+If we let $M\rightarrow\infty$ then we achieve asymptotic equality to the true variance estimate.
 
-	If models are uncongenial, we use Meng's rules.
+	Xia and Meng (2017) state that ''When the imputation model class and the (embedded) analyst’s model class
+	differ, the behavior of Rubin’s rules becomes very complicated, capable of producing inconsistent variance estimators, a matter that has received recurrent
+	criticisms''.
+They argue that in the uncongenial case , which is often the rule rather than the exception, doubling Rubin's variance measure will never underestimate the true variance.
+This holds for any degree of uncongeniality, and provides a surprisingly simple solution to the problem of uncongeniality.
+For Univariate data, Xia and Meng (2017) suggest replacing $T$ with $\tilde{T}= (\sqrt{\bar{Var}}+\sqrt{B})^{2}+\frac{1}{M}B_$.
+These methods may be viewed as highly conservative, but in the context of MI where uncongeniality between analysts and imputer is unknown and potentially large, it pays to be as insured as possible against invalid inference.
+However, even if these methods are not used and the models are uncongenial, we note that Meng (1994) has shown that MI is still superior to incomplete data procedures.
 
@@
 
+	%Improper specification of the imputation model causes uncongeniality and introduces bias in the data that can distort inference.
+
 	Even Rubin (1996) admits that MI
@@
 As long as a potential user is aware of the potential pitfalls, MI is the superior method for dealing with missing data.
+
+	%Needs more pitfalls and criticism
 
@@
 	MI can be used to perform sensitivity analysis by varying the imputation model.
+
+	%Needs to be more comprehensive.
 
@@
 	Citation example:
+	Self efficiency.
+
+	\subsection{Final notes}
+	This paper has gone into great detail on MI and missing data in general.
+However, a rigorous comparison between likelihood based methods and MI has not been performed in the literature to the same extent as this paper.
+Espceially in the case of uncongenial imputation models, one would think likelihood based methods might show some promise.
 	\cite[P.10]{grund_pooling_2016}.

### 2020-03-05 experiments.Rmd
Minor additions to code

@@
 library(mvtnorm)
+library(MASS)
+```
+
+```{r}
+#Setting number of entries
+n = 1000
 ```
@@
 #Setting vals for multivariate continous vars
-mu = 8
-sigma = 2
-help(mvtnorm)
+mu = replicate(14,0)
+tez = matrix(c(1, replicate(10,0)), c(2, replicate(9, 0), 3), nrow=2, ncol=11)
+tez
+#Variables only correlated with Y
+sigmaNoCrossCorr = matrix(c(c(1,  0.4, 0.6, 0.8, 0.1, 0.1, 0.1, 0.2, 0.2, 0.2, 0.4, 0.4, 0.8, 0.8),
+          c(0.4, 1, replicate(12, 0)),
+          c(0.6, 0,  1, replicate(11,0)),
+          c(0.8, 0, 0, 1, replicate(10,0)),
+          c(0.1, replicate(3,0), 1, replicate(9,0)),
+          c(0.1, replicate(4,0), 1, replicate(8,0)),
+          c(0.1, replicate(5,0), 1, replicate(7,0)),
+          c(0.2, replicate(6,0), 1, replicate(6,0)),
+          c(0.2, replicate(7,0), 1, replicate(5,0)),
+          c(0.2, replicate(8,0), 1, replicate(4,0)),
+          c(0.4, replicate(9,0), 1, replicate(3,0)),
+          c(0.4, replicate(10,0), 1, replicate(2,0)),
+          c(0.8, replicate(11,0), 1, 0),
+          c(0.8, replicate(12,0), 1)), nrow=14, ncol=14)
+sigmaNoCrossCorr
+
+#variables correlated with eachother, still under construction
+sigmaCrossCorr = c(c(1,  0.4, 0.6, 0.8, 0.1, 0.1, 0.1, 0.2, 0.2, 0.2, 0.4, 0.4, 0.8, 0.8),
+          c(0.4, 1, replicate(12, 0)),
+          c(0.6, 0,  1, replicate(11,0)),
+          c(0.8, 0, 0, 1, replicate(10,0)),
+          c(0.1, replicate(3,0), 1, replicate(9,0)),
+          c(0.1, replicate(4,0), 1, replicate(8,0)),
+          c(0.1, replicate(5,0), 1, replicate(7,0)),
+          c(0.2, replicate(6,0), 1, replicate(6,0)),
+          c(0.2, replicate(7,0), 1, replicate(5,0)),
+          c(0.2, replicate(8,0), 1, replicate(4,0)),
+          c(0.4, replicate(90,0), 1, replicate(3,0)),
+          c(0.4, replicate(10,0), 1, replicate(2,0)),
+          c(0.8, replicate(11,0), 1, 0),
+          c(0.8, replicate(12,0), 1))
+mvnData = mvrnorm(n = n, mu, sigmaNoCrossCorr, tol = 1e-14)
 ```
@@
 #Generating binary data
-n = 1000
+
 binary_categorical_vector = round(runif(n, min = 0, max = 1))

### 2020-03-15 Draft1.tex
Minor changes

@@
 	\subsection{Congeniality: when the imputer's and analysts models differ}
+	Congeniality is an important concept in imputation and has a big impact on the statistical validity of MI.
+In short, uncongeniality arises when the analysis model differs substantially from the imputation model.
+There are many reasons for such difference, for example, the imputer might wish to include as many variables as possible to achieve an accurate imputation, while the analyst might be interested in only a subset of the covariates and the relationship between these.
+However, as Meng [1994] states, ''If the imputer's model is reasonably accu-
+	rate, then following the multiple-imputation recipe
+	prevents the analyst from producing inferences with
+	serious nonresponse biases''.
+Meng goes on to show that even with an uncongenial imputation model, inferences performed on a MI dataset still yield better inference than the best possible incomplete data inference.
+
 	As mentioned in the introduction, modern data analysis' multiple phases mean that the assumptions underlying each phase and the relationship between the phases is an important object of study.
@@
 Third is when one is conducting sensitivity analysis, which necessarily requires the testing of different models, some of which will be uncongenial.
-
-	Congeniality is an important concept in imputation and has a big impact on the statistical validity of MI.
-In short, uncongeniality arises when the analysis procedure differs substantially from the imputation model.
-There are many reasons for such difference, for example, the imputer might wish to include as many variables as possible to achieve an accurate imputation, while the analyst might be interested in only a subset of the covariates and the relationship between these.
-However, as Meng [1994] states, ''If the imputer's model is reasonably accu-
-	rate, then following the multiple-imputation recipe
-	prevents the analyst from producing inferences with
-	serious nonresponse biases''.
-Meng goes on to show that even with an uncongenial imputation model, inferences performed on a MI dataset still yield better inference than the best possible incomplete data inference.
 
@@
 
-	\textbf{Last Observation Carried Forward (LOCF):x}
+	\textbf{Last Observation Carried Forward (LOCF):}
 	This method is mostly used in longitudinal studies where individuals drop out, withdraw, or are lost over the course of the studies.
@@
 This holds for any degree of uncongeniality, and provides a surprisingly simple solution to the problem of uncongeniality.
-For Univariate data, Xia and Meng (2017) suggest replacing $T$ with $\tilde{T}= (\sqrt{\bar{Var}}+\sqrt{B})^{2}+\frac{1}{M}B_$.
+For Univariate data, Xia and Meng (2017) suggest replacing $T$ with $\tilde{T}= (\sqrt{\bar{Var}}+\sqrt{B})^{2}+\frac{1}{M}B$.
 These methods may be viewed as highly conservative, but in the context of MI where uncongeniality between analysts and imputer is unknown and potentially large, it pays to be as insured as possible against invalid inference.

### 2020-06-11 Draft1.tex
Finished

--- 2020-03-03 Draft1.tex
@@
 \usepackage{graphicx}
-
-\setlength{\parindent}{4em}
-\setlength{\parskip}{1em}
-
-\title{Missing data: a simple introduction}
-\author{Marius Dioli}
-\date{February 2019}
+\usepackage{amsmath}
+\usepackage{amsthm}
+\usepackage{xfrac}
+\usepackage{float}
+\usepackage{amssymb}
+\usepackage[font={small}]{caption}
+\usepackage{csquotes}
+\usepackage{subfig}
+\usepackage{booktabs}
+\usepackage{tabularx}
+\usepackage{titling}
+\newcommand{\subtitle}[1]{%
+	\posttitle{%
+		\par\end{center}
+	\begin{center}\large#1\end{center}
+	\vskip0.5em}%
+}
+\usepackage{hyperref}
+\captionsetup[figure]{labelfont=bf,textfont={it}}
+\captionsetup[table]{position=above,labelfont=bf,textfont={it}}
+\graphicspath{{figures/}}
 
@@
 
+	\section*{Abstract}
+	Because of the centrality of data analysis in many human endeavors from science to industry, methods of dealing with faults and errors in data are an important area of study.
+One such error is the problem of missing data.
+This thesis provides the reader with a comprehensive introduction to the field of missing data.
+It describes the concept of missingness mechanisms and how they might impact inference.
+Imputation is evaluated and found to be a statistically valid technique for addressing missing data in the context of these missingness mechanisms.
+Various common and intuitive imputation methods such as Complete Case Analysis and Mean Imputation are analysed.
+Evidence from theory and literature is presented and lead to the conclusion that Multiple Imputation (MI) is the superior method.
+However, this is contradicted by experimental results, suggesting further study is required to uncover the source of this disagreement.
+
 	\section{Introduction}
-	... ''eng and Romero (2003), Xie and Meng (2012)). In general,
-	the problem of multiphase inference can be formulated as one of missing data. However, in the
-	multiphase setting, missingness arises from the preprocessing choices made, not a probabilis
-	tic response mechanism. Thus, we can leverage the mathematical and computational methods
-	of this literature, but many of its conceptual tools need to be modified. Multiple imputation ad
-	dresses many of the same issues as multiphase inference and is indeed a special case of the
-	latter. Concepts such as congeniality between imputation and analysis models and self-efficiency
-	(Meng ( 1994)) have natural analogues and roles to play in the analysis of multiphase inference
-	problems.''
-
-	In this paper we will focus on the Multiple Imputation in the context of correcting for missing data before analysis.
-Blocker and Meng argue that general preprosessing can be evaluated in the missing data paradigm, but this is outside the scope of this paper.
-We refer readers to Block and Meng for an in depth consideration of the subject.
-
-	The public database is perhaps the context where MI is most useful.
-These are often analyzed by multiple users with varying degrees of statistical knowledge and resources.
-All public databases have missing data, and typically users do not have the knowledge to deal with missing data satisfactoraly.
-Even if they do, the database constructors often have more information about the missing data than the end user, and so can better compensate for this.
-	%TODO Introduce the concept of multiphase inference. Reference meng2017 and Blocker and meng
-	If one is to do anything of import in science, one has to manipulate and analyze data.
-However, due to a variety of factors, data is often flawed.
-One major flaw one can have, and the topic of this thesis, is missing data.
-The consequence of missing data is always a reduction in efficiency and may also lead to bias.
-
-	//A variety of ad hoc approaches are commonly used to deal with missing data, and many of the techniques described in this thesis are either//
-
-	//This thesis will approach the problem of missing data from the bayesian perspective and with the goal of imputing or removing data.
-Other methods motivated from the frequentist perspective such as likelihood based methods will be touched on breifly as alternatives.
-The strength of the bayesian approach is that it provides a principled way to account for uncertainty about the missingness mechanism and the true distribution of the full data.
-//
-
-	To quote MENG: ''imputation is not (merely) a compu-
-	tational tool but rather a mode of inference, which al-
-	lows hierarchical and sequential input of assessment
-	and information'' [Multiple-Imputation Inferences with Uncongenial Sources of Input]
-
-	.. We will measure its efficacy on synthetic data and real data before comparing it with an alternate method that is gaining in popularity, machine learning.
-
-	This paper is by no means comprehensive, but strives to give the reader a proper introduction to the problem of missing data and the technique of multiple imputation.
-	// fyldig
-
-	As detailed in Xie and Meng [DISSECTING MULTIPLE IMPUTATION FROM], much of modern data analysis is done in a multi-phase paradigm, where the distinct and sequential phases data acquisition, data pre-processing, and data analysis, can have different or even contradictory assumptions.
-
-	Our objective with multiple imputation, as stated in Rubin(1996), is: assuming that ultimate user's complete-data analysis is stat
-
-	Although our main objective with MI is inference, we will test its ability to improve performance in prediction and classification contexts as well.
-
-	The goal of this paper is to give an overview of missingness theory and multiple imputation such that as many as possible can then apply MI well.
+	Wherever there is data there tends to be missing data, and missing data can distort or invalidate the conclusions drawn from data.
+In science and academia data analysis is fundamental to progress, but gathering data in the real world can prove tricky.
+Take clinical trials of a drug, where dropout can be a serious obstacle to figuring out whether a drug is worth producing or not.
+Similarly, from tech to traditional heavy industry, data analysis allows for efficiency gains and innovation.
+However, things from faulty sensors to privacy laws introduce missing data.
+In this thesis, we shall give an overview of the mathematical theory concerning missing data, as well as some techniques for dealing with the problem.
+We will first describe common and intuitive solutions, so called \enquote{ad-hoc} solutions, before proceeding to a statistically principled and effective solution: Multiple Imputation (MI).
+We will approach the problem of missing data from the Bayesian perspective.
+Other methods motivated from the frequentist perspective, such as likelihood based methods, are available for dealing with missing data, but shall not be explored here.
+For the sake of brevity and cohesiveness, some topics like congeniality have only been covered in light detail.
+We refer the reader to Meng [1994], Molenberghs et.
+al.
+[2015], and Little and Rubin [2019] for a more in-depth explanation of missing data and non-imputation based techniques.
 
 	\section{Terminology and basic missingness theory}
-	In the following section I will mostly adhere to the notation of Rubin in (citation), as it is both well established and quite general.
-However, I will adapt it slightly by using the observed/missing notation in (missing data) for readability, as it is easier for an unfamiliar audience and non-specialists to understand.
+	This section will mostly adhere to the notation of Rubin in \cite{Little2019}, as it is both well established and quite general.
 
 	\subsection{What is missingness?}
-	When we say a variable is missing we mean that we have not observed it.
-The variable still has a value, but this value is hidden from us for some reason.
-What this reason might be will be described in detail later in this section.
-By treating missingness as a probabalistic phenomenon, statistically rigorous tools for dealing with it become available.
+	When we say an observation is missing, we mean that we have not observed it.
+The missing observation still has a value, but this value is hidden from us for some reason.
+The mechanism behind this missingness will be described in detail later in this section.
+By treating missingness as a probabilistic phenomenon, statistically rigorous tools for dealing with it become available.
 
-	Let $Y=(y_{ij})$ be a $n*k$ complete data matrix (i.e.
-with no missing values), with the ith row $y_{i}=(y_{i1}, ..., y_{ik})$ where $y_{ij}$ is the value of variable $Y_{j}$ for unit $i$.
-Let also the missingness indicator matrix $M = (m_{ij})$ be an $nxk$ such that $m_{ij} = 1$ if $y_{ij}$ is missing and $m_{ij} = 0$ if $y_{ij}$ is observed.
-Both $Y$ and $M$ are random variables, and we assume for simplicity that the rows $(y_{i}, m_{i})$ are independent and identically distributed over i.
+	Let the complete data matrix (i.e.
+with no missing values) $Y=(y_{ij})$ be an $n \times k$ matrix, with the $i^{\text{th}}$ row $y_{i}=(y_{i1}, ..., y_{ik})$ where $y_{ij}$ is the value of variable $Y_{j}$ for observation $i$.
+Let also the missingness indicator matrix $M = (m_{ij})$ be an $n \times k$ matrix such that $m_{ij} = 1$ if $y_{ij}$ is missing and $m_{ij} = 0$ if $y_{ij}$ is observed.
+Both $Y$ and $M$ are random variables, and we assume for simplicity that the rows $(y_{i}, m_{i})$ are independent and identically distributed over $i$.
 It is possible to have more than two possible values for each entry of the $M$, for example when we wish to differentiate between different kinds of non-response.
-However, we will not explore this in this paper.
-Now, given $m_{i}$, we can partition $y_{i}$ into $y^{o}_{i}$ and $y^{m}_{i}$, corresponding to the components that are observed and missing respectively (where the corresponding entry in $m_{i}$ is 0 or 1 respectively).
+However, we will not explore this in this thesis.
+Now, given $m_{i}$, we can partition $y_{i}$ into $y^{o}_{i}$ and $y^{m}_{i}$, corresponding to the components that are observed and missing respectively.
 
@@
 $f_{M|Y}(m_{i}|y_{i}, \theta)$, where $\theta$ denotes unknown parameters.
-Stated differently the missingness mechanism is the probability of data being missing given the values of Y, missing and observed.
+Stated differently, the missingness mechanism characterizes the probability of data being missing given the values observed values of $Y$ and the information that the missing values of $Y$ are missing.
 
-	The missingness pattern in $Y$ does not matter for the following definitions, and can be any of the following: *(Figure from page 9 Rubin)* 
-	However, the missingness pattern will affect the efficacy of methods used to deal with missingness, and we shall explore this in subsequent sections.
+	We distinguish between three different missingness mechanisms by considering how $M$ is related to $Y$ in order of decreasing strictness of assumptions:
 
-	We distinguish between three different missingness mechanisms by considering how $M$ is related to $Y$: Missing Completely at Random (MCAR), Missing at Random (MAR), Missing Not at Random (MNAR).
-These can be roughly thought of as being in order of strictness from strictest (MCAR), to least strict (MNAR).
-Rubin(1976) [missing data book] introduced these missingness mechanisms, and they determine which techniques are appropriate to use on a given problem.
-**These mechanisms are rarely the case in practice, but it will be shown later that assuming them is reasonable in many circumstances.
-**
+	\begin{enumerate}
+		\item Missing Completely at Random (MCAR)
+		\item Missing at Random (MAR)
+		\item Missing Not at Random (MNAR)
+	\end{enumerate}
 
-	\textbf{Missing Completely At Random (MCAR)}:
-	This kind of missingness occurs when, $\forall i$ and any distinct values $y_{i}, y^{\ast}_{i}$ in the sample space of Y, $f_{M|Y}(m_{i}|y_{i}, \theta) = f_{M|Y}(m_{i}|y^{\ast}_{i}, \phi)$.
-In other words, it means that the probability of missingness is independent of both the observed and the missing data, and we can view the observed data as a random sample of the complete data.
+	 Rubin [1976] introduced these missingness mechanisms, and they determine which techniques are appropriate to use on a given problem.
+In practice, they are rarely exactly true, but it will be shown later that assuming them is reasonable in many circumstances.
 
-	\textbf{Missing At Random (MAR)}:
-	In this case, $\forall i$ and any distinct values $y^{m}_{i}, y^{m\ast}_{i}$ in the sample space of $y^{m}_{i}$, $f_{M|Y}(m_{i}|y^{o}_{i}, y^{m}_{i}, \theta) = f_{M|Y}(m_{i}|y^{o}_{i}, y^{m\ast}_{i}, \theta)$.
+	\subsubsection{Missing Completely At Random}
+	 MCAR missingness occurs when, for all and any distinct values $y_{i}$ and $ y^{\ast}_{i}$ in the sample space of $Y$, $$f_{M|Y}(m_{i}|y_{i}, \theta) = f_{M|Y}(m_{i}|y^{\ast}_{i}, \phi), \quad \exists \theta, \phi$$ where $\theta$ and $\phi$ denote unknown parameters.
+In other words, the probability of missingness is independent of both the observed and the missing data.
+This means that we can view the observed data as a random sample of the complete data.
+
+	\subsubsection{Missing At Random}
+	In this case, for all i and any distinct values $y^{m}_{i}, y^{m\ast}_{i}$ in the sample space of $Y^{m}_{i}$, we have that the following equation holds: $$f_{M|Y}(m_{i}|y^{o}_{i}, y^{m}_{i}, \theta) = f_{M|Y}(m_{i}|y^{o}_{i}, y^{m\ast}_{i}, \phi), \qquad \exists \theta, \phi$$
 	This means that the probability of missingness, conditioned on the observed data, is independent of the missing data.
-Assuming MAR is less restrictive than MCAR, and may be considered a more plausible assumption about missing data in many contexts.
+Assuming MAR is less restrictive than MCAR, and may be considered a more plausible assumption in a real world situation.
 
-	\textbf{Missing Not At Random (MNAR)}:
+	\subsubsection{Missing Not At Random}
 	If the above equation does not hold for some $i$ and some $y^{m}_{i}, y^{m\ast}_{i}$, then the missingness mechanism is called missing not at random (MNAR).
-This means that the probability of missingness is dependent on the missing data even after conditioning on the observed data.
+In this situation, the probability of missingness is dependent on the missing data even after conditioning on the observed data.
+Put differently, this means that the underlying values of the missing data influence the probability of data being missing.
+Thus the the missingness mechanism cannot be characterised from the available data in the way that MAR can.
 
@@
 	Suppose you are doing a study in which you want to investigate people's sleep patterns.
-You send N people a survey asking them for the average amount of sleep they got each night in 2018.
-You send a follow-up survey for 2019.
-Table 1 shows simulated data for N = 15 people drawn from a bivariate normal distribution with $\mu_{x} = \mu_{y} = 7.5$, $\sigma_{x} = \sigma_{y} = 2$, and a correlation $\rho = 0.4$.
-The first two columns show the complete data for X and Y, while the other three columns show the values for 2019 after imposing missingness by three methods.
+You send $N$ people a survey asking them for the average amount of sleep they got each night in 2018.
+A year later, you decide to send a follow-up survey for 2019.
+We consider 2018 and 2019 to be stochastic variables.
+Table 1 shows a simulation of such a study.
 
-	\begin{figure}
-		\centering
-		\includegraphics[width=\linewidth]{missingness}
-		\caption{Different missingness mechanisms}
-		\label{fig:miss1}
-	\end{figure}
+	\begin{table}[]
+		\begin{tabular}{ccccc}
+			\multicolumn{1}{l}{2018} & \multicolumn{1}{l}{2019 (Complete)} & \multicolumn{1}{l}{2019 (MCAR)} & \multicolumn{1}{l}{2019 (MAR)} & \multicolumn{1}{l}{2019 (MNAR)} \\ \hline
+			8.9                      & 7.4                                 & 7.4                             & 7.4                            & NA                              \\
+			9.4                      & 7.3                                 & 7.3                             & 7.3                            & NA                              \\
+			8.4                      & 7.9                                 & NA                              & 7.9                            & NA                              \\
+			5.3                      & 6                                   & NA                              & NA                             & NA                              \\
+			11.1                     & 7.4                                 & 7.4                             & 7.4                            & NA                              \\
+			8.4                      & 10.7                                & NA                              & 10.7                           & 10.7                            \\
+			6.4                      & 8.8                                 & NA                              & NA                             & 8.8                             \\
+			5.6                      & 3.5                                 & NA                              & NA                             & NA                              \\
+			5.6                      & 7.1                                 & 7.1                             & NA                             & NA                              \\
+			8.2                      & 6.9                                 & 6.9                             & 6.9                            & NA                              \\
+			6.6                      & 5.4                                 & NA                              & NA                             & NA                              \\
+			6.6                      & 6.7                                 & NA                              & NA                             & NA                              \\
+			8.6                      & 9.2                                 & NA                              & 9.2                            & 9.2                             \\
+			7.7                      & 6.1                                 & 6.1                             & NA                             & NA                              \\
+			11.2                     & 8                                   & NA                              & 8                              & 8                              
+		\end{tabular}
+	\caption{\label{tab:tab1}Simulated data for $N = 15$ individuals drawn from a bivariate normal distribution with $\mu_{2018} = \mu_{2019} = 7.5$, $\sigma_{2018} = \sigma_{2019} = 2$, and a correlation $\rho = 0.3$.
+The first two columns show the complete data for 2018 and 2019, while the other three columns show the values for 2019 after imposing missingness by the three methods discussed earlier.}
+	\end{table}
 
@@
 a random subset of those who answered the 2018 survey also answered the 2019 one.
-Things get more interesting in the next two columns.
-Suppose you are only interested in those who've slept less than 7.5 hours on average, so you only send a follow-up survey to those who reported less than 7.5 hours of sleep for 2018 (2018<7.5).
+Now, suppose that the amount of sleep people got in 2018 affected how likely they were to answer the 2019 survey.
+Those who slept little in 2018 were less likely to answer in 2019, and conversely with those who slept more.
 This is MAR, since the missingness of the 2019 entries depends on the values for 2018.
-For the final column, suppose you sent out your 2019 survey to all the individuals who answered the 2018 one, but include a line stating "Only answer this survey if you slept less than 7.5 hours on average in 2019" (2019 < 7.5).
-This is MNAR, since the missingness of the 2019 entries depends on the values for those entries.
-Remember, the full dataset exists, but it has been censored due to some mechanism, in this case through your own design.
+For the final column, suppose our earlier conjecture is instead true for 2019: those who slept little in 2019 were less likely to answer the survey in that year.
+This is an example of MNAR, since the missingness of the 2019 entries depends on the values for those entries.
+Remember, the full dataset exists, but it has been hidden from us due to some mechanism.
 
@@
 
-	In controlled environments, like the one from our example, we can guarantee that data are MAR since we control the censoring of the data (in this case by placing a cutoff between 2018 and 2019).
-However, it is rare that we have this much control over our data, and a compelling reason to censor it so.
-In most real world scenarios we only have the data at hand to inform us about the missingness mechanism.
-Unfortunately, we cannot infer the missingness mechanism from data.
-We can check the validity of MCAR against MAR, but only under the unverifiable assumption that the data is not MNAR.
-Neither can we use the data at hand to support or refute on specific MNAR mechanism over another.
-The literature stresses the importance of sensitivity analyses in this case, checking the sensitivity of our inference under a variety of plausible assumptions.
-Finally MAR cannot be checked empirically against MNAR  (Cite from page 8, 9, and 10 in missing data).
-We therefore have to assume the missingness mechanism, most often MAR, and there are strong arguments for this.
-First and foremost is that assuming MAR yields good empirical results, which will be shown in later sections.
-%TODO Add more reasons for assuming MAR
+	In most real world scenarios, we only have the data at hand to inform us about the missingness mechanism.
+Unfortunately, we cannot infer the missingness mechanism from data \cite[P.9]{Molenberghs2015}.
+We can check the validity of MCAR against MAR, but only under the unverifiable assumption that the data is not MNAR \cite[P.8]{Molenberghs2015}.
+Neither can we use the data at hand to support or refute one specific MNAR mechanism over another \cite[P.10]{Molenberghs2015}.
+Finally, MAR cannot be checked empirically against MNAR.
+We therefore have to assume the missingness mechanism in practice.
+Most often it is beneficial to assume MAR, as it tends to yield good empirical results even when the assumption does not completely hold.
 
@@
 	\subsection{What is imputation?}
-	Imputation refers to substituting or ''filling in'' of missing values with plausible estimates.The main strength of imputation is that it does not change the nature of the subsequent analysis.
-It makes it much easier for the data processor and analyst to work together since the processor's techniques won't interfere with the analysts'.
-This is, at least, the goal of imputation.
-In certain situation the information available to the imputer might differ to that available to the analyst.
-In such cases difficulties arise.
+	Imputation refers to the substituting, or \enquote{filling in}, of missing values with plausible estimates.
+Often there much information available in the data at hand, and utilising this can lead to better performance.
+The main strength of imputation is that it does not require us to change our desired analysis technique.
+In certain situations, a researcher might have more data than they intend to include in their analysis model.
+This extra data can be used in the imputation process to produce a better imputation, but can introduce bias or variance in parameter estimates.
 The concept of congeniality, which we will define shortly, helps us overcome some of these difficulties.
-Another very practical benefit, and a strong argument for its use, is the performance gains to be had from using it.
-This will become clear in the experiments section.
+Another very practical benefit, and a strong argument for the use of imputation, is the performance gains to be had from using it.
+Examples can be seen in \cite{madley-dowd_proportion_2019}, \cite{Schafer2002}, and \cite{Sterne2009}, as well as in the experiments section.
 
-	The more advanced model based approaches of this paper uses both the Bayesian and frequentist paradigms in complementary ways.
-The Bayesian approach creates procedures, while the frequentist one evalueates them.
-This is because the bayesian paradigm allows us to effectively model the unknown given the known, and makes assumptions explicit.
-The frequentist paradigm is the one which is by far most common when doing inference, so it makes sense that it is under that paradigm performance matters.
+	The model based approaches of this thesis use both the Bayesian and frequentist paradigms in complementary ways.
+The Bayesian approach is used to create the imputation procedure, while the frequentist one evaluates their results.
+The Bayesian paradigm allows us to effectively model the unknown given the known, and makes assumptions explicit, both highly desirable when imputing missing data.
+The frequentist paradigm is the one which is by far most common when doing inference, so it is natural that it is under that paradigm we evaluate performance.
 
 	\subsection{What is valid inference?}
-	Techniques are based on the bayesian paradigm, but inference is often done in the frequentist one.
+	An essential question when using imputation is if, and how, it might distort inference.
+However, to answer this question we need to have an idea of what valid inference is.
+A valid inference, according to Rubin \cite{rubin_multiple_1996}, is one in which nominal significance levels (denoted by $\alpha$ and often set at $\alpha = 0.05$) describe characteristics of a distribution for confidence interval estimation and hypothesis testing.
+This can be further refined into \enquote{randomization validity} and \enquote{confidence validity}.
+The former Rubin describes as \enquote{actual interval coverage = nominal interval coverage}, for interval estimates, and  \enquote{actual rejection rate = nominal rejection rate}, for hypothesis tests.
+Randomization validity is of course desirable, but not always achievable, and so Rubin introduces confidence validity as a looser criterion.
+It is defined as \enquote{actual interval coverage $\geq$ nominal interval coverage}, for interval estimates, and \enquote{actual rejection rate $\geq$ nominal rejection rate}, for hypothesis testing.
+It is this last version of inferential validity that we shall be referring to whenever we talk about valid inference.
+It has been shown \cite{Xie2017}, \cite{Meng1994}, \cite{rubin_multiple_1996} that Bayesian models like the Multiple Imputation (MI), which we will describe in detail in a later section, yield valid inference.
 
-	Rubin (1996) defines the term scientific estimand as the objects we wish to draw inference about. He defines them thusly: "... a quantity of scientific
-	interest that can be calculated in the population and does
-	not change its value depending on the data collection de-
-	sign used to measure it (i.e., it does not vary with sample
-	size and survey design, or the number of nonrespondents,
-	or follow-up effort. Sci-
-	entific estimands include population means, variances, cor-
-	relations, factor loadings, regression coefficients, and these
-	quantities within strata or domains, but exclude the sam-
-	pling variance of a sample mean under a particular sampling
-	plan and the expectation of the complete-data sample mean
-	when missing values are filled in with zero or the observed sample mean.
-".
-Rubin goes on to describe statistical validity for scientific estimands as the following: " point estimation must be approximately unbi-
-	ased for the scientific estimands, averaging over the sam-
-	pling and the posited nonresponse mechanisms (e.g., filling
-	in zeros or means is not generally acceptable)." and "in-
-	terval estimation and hypothesis testing must be valid in
-	the sense that nominal levels describe operating character-
-	istics over sampling and posited response mechanisms.".
-The latter criteria can be further refined into "randomization validity" and "confidence validity".
-The first is that, according to Rubin (1996),  "randomization validity means that, for interval estimates,
-	"actual interval coverage = nominal interval coverage," and
-	for tests of hypotheses, "actual rejection rate = nominal re-
-	jection rate."  Randomization validity is of course desireable, but not always achievable, and so Rubin introduces confidence validity as a looser criterion and defines it as "hat for interval estimates, "ac-
-	tual interval coverage >= nominal interval coverage," and for
-	tests of hypotheses, "actual rejection rate =< nominal rejec-
-	tion rate."" We therefore proceed in our analysis of methods using these definitions when evaluating the inferential validity of imputation methods.
+	\subsection{Congeniality: when the imputation and analysis models differ}
+	Congeniality is an important concept in imputation and has a big impact on the statistical validity of imputation.
+In short, when the analysis model differs substantially from the imputation model we say that they are uncongenial.
+Such differences might cause the subsequent analysis to be biased, or in the worst case, invalid.
+There are many reasons for such difference, for example, one might wish to include as many variables as possible in the imputation model to achieve an accurate imputation, while at the same time only being interested in only a subset of the covariates and the relationship between these.
+We will not go into detail about what specifically constitutes a \enquote{substantial difference}, and refer to \cite{Meng1994} for a detailed description and explanation.
+However, an important finding by Meng is that even with an uncongenial imputation model, MI still outperforms methods that ignore missing data.
+This will be borne out in the experiments section.
 
-	MI is motivated under the Bayesian paradigm.
-This causes some (Reference Fay and stuff from MENG) to question whether inference under MI is valid at all.
-To address this Meng [1994] developed the concept of congeniality, which we will go into more detail on soon.
-
-	\subsection{Congeniality: when the imputer's and analysts models differ}
-	As mentioned in the introduction, modern data analysis' multiple phases mean that the assumptions underlying each phase and the relationship between the phases is an important object of study.
-Meng's [meng 1994] concept of congeniality provides a theoretical framework from which to evaluate these differences and most of this section is derived from his paper.
-
-	First let us define some of the basic notation that Meng (1994) uses to define congeniality.
-We will modify it slightly to fit with the notation of this paper and since Meng's paper was written in the context of surveys.
-An analyst's complete-data procedure can be summarised by $P_{com} = {\hat{Q}(Y, M), U(Y, M)}$ where $\hat{Q}(Y, M)$ is an estimator of $Q(Y, M)$, an uknown quantity of interest, with $U(Y, M)$ being an associated variance estimator.
-Likewise, an analyst's incomplete data procedure is denoted by $P_{obs} = {\hat{Q}(Y^{o}, M), U(Y{o}, M)}$.
-Meng goes on to define congeniality in the following manner: "A Bayesian model f is said to be congenial to the analysis procedure $P = {P_{obs};P_{com}}$ for a given ${Y^{o}, M}$ if the following hold:
-
-	1) The posterior mean and variance of $Q$ under
-	f given the incomplete data are asymptotically the
-	same as the estimate and variance from the analyst's
-	incomplete-data procedure , $P_{obs}$ that is, $$[[\hat{Q}(Y^{o}, M), U(Y^{o}, M)]\simeq [E_{f}[Q|{Y^{o}, M}], Var_{f}[Q|{Y^{o}, M}]]$$
-
-	2)  The posterior mean and variance of $Q$ under
-	f given the complete data are asymptotically the
-	same as the estimate and variance from the analyst's
-	complete-data procedure , $P_{com}$ that is, $$[[\hat{Q}(Y), U(Y)]\simeq [E_{f}[Q|{Y}], Var_{f}[Q|{Y}]]\qquad \forall Y\textrm{with}Y^{o}\textrm{fixed (i.e., conditioned upon)}$$
-
-	Additionally, let $g(Y^{m}|{Y^{o}, M}, A)$ be the imputation model, where A represents possible additional data the imputer has access to.
-Meng continues by stating ''The analysis procedure $P$ is said to be congenial to the imputation model $g(Y^{m}|{Y^{o}, M}, A)$ if one can find an f such that 1) f is congenial to $P$ and 2) the posterior predictive density for $Y^{m}$ derivd under f is identical to the imputation model $$f(Y{m}|{Y^{o}, M}) = g(Y{m}|{Y^{o}, M}, A) \forall Y^{m}$$
-
-	Therefore if no such congenial f exists, then$P$ is said to be uncongenial to g.
-Notice that congeniality is defined with respect to a particular analysis procedure, meaning that an analyst can perform two types of analyses where one is congenial and the other is not.
-There are three main scenarios in which uncongeniality occurs in the context of missing data and MI.
-The first is when the imputation model is largely unknown to the analyst, often the case with public databases.
-Second is when the different needs of imputation and analysis require different models.
-Third is when one is conducting sensitivity analysis, which necessarily requires the testing of different models, some of which will be uncongenial.
-
-	''When the imputation model class and the (embedded) analyst’s model class
-	differ, the behavior of Rubin’s rules becomes very complicated, capable of producing inconsistent variance estimators, a matter that has received recurrent
-	criticisms'' [Meng 2017]
-
-	Congeniality is an important concept in imputation and has a big impact on the statistical validity of MI.
-In short, uncongeniality arises when the analysis procedure differs substantially from the imputation model.
-There are many reasons for such difference, for example, the imputer might wish to include as many variables as possible to achieve an accurate imputation, while the analyst might be interested in only a subset of the covariates and the relationship between these.
-However, as Meng [1994] states, ''If the imputer's model is reasonably accu-
-	rate, then following the multiple-imputation recipe
-	prevents the analyst from producing inferences with
-	serious nonresponse biases''
-
-	\subsection{Techniques of interest}
+	\section{Techniques of interest}
 	To properly evaluate the performance of MI, we should compare it to other similar and commonly used techniques, i.e.
@@
 Second, does the method give us appropriate measures of precision?
-How the following methods answer these questions is how we will decide the superiority of one method over another.
+How the following methods answer these questions will determine which methods are preferable.
+For a more in-depth description and examples of these methods we refer the reader to \cite{Molenberghs2015}.
 
-	There are two other competing methods which we should touch on briefly before continuing.
-Weighting procedures and likelihood-based methods are ...
-alternatives to multiple imputation.
-However, there is reason to prefer MI as the default options over these for a few reasons.
-First that, given congeniality, MI tends to outperform likelihood-based methods for small sample sizes, while delivering similar performance for large sample sizes \cite{schafer_multiple_2016}.
+	There is another competing set of methods which we should touch on briefly before continuing: likelihood-based methods.
+These methods focus on maximising the likelihood function directly.
+Under MAR, the missing data can be ignored in calculations, leading to a much more tractable problem.
+However, given congeniality, more advanced imputation methods like multiple imputation tend to outperform likelihood-based methods for small sample sizes, while delivering similar performance for large sample sizes \cite{schafer_multiple_2016}.
 
 	\subsection{Ad hoc methods}
-	This section is mostly adapted from Carpenter and Kenward [Ch2.5 missing data].
-Although Ad hoc is pejorative in certain circumstances, it is used here to indicate that the methods below are motivated by convenience rather than any methodological concern.
+	This section is mostly adapted from Chapter 2 of \cite{Molenberghs2015} Although ad hoc is pejorative in certain circumstances, it is used here to indicate that the methods below are motivated by convenience rather than any methodological concern.
 
-	\textbf{Case deletion, also called complete case analysis (CCA):}
-	Probably the most common technique in use for dealing with missing data, CCA involves restricting analysis to entries with complete data.
+	\subsubsection{Case deletion, also called Complete Case Analysis}
+	Probably the most common technique in use for dealing with missing data, Complete Case Analysis (CCA) involves restricting analysis to entries with complete data.
 For those familiar with programming, a typical example would be to remove all rows from a dataframe with NaN entries.
 The first and most obvious problem with this approach is that it drastically reduces the amount of useable data.
-This holds an intuitive appeal, as it is both exceedingly simple to implement, and seem to have the regrettable but sometimes minor effect of lower accuracy.
-However, depending on the missingness mechanism, this might be wrong, and can actually introduce bias into the data.
+CCA holds an intuitive appeal, as it is both exceedingly simple to implement, and with the only seeming downside being a loss in accuracy.
+However, depending on the missingness mechanism, CCA can actually introduce bias into an analysis.
 For example, suppose you want to find the mean and standard deviation in the height of a population.
@@
 An implicit assumption, therefore, of CCA is that the observed data is a random sample of the underlying distribution.
-This means that CCA implicitly assumes MCAR, and we see in [schafer and graham] that we get unbiased inference in this scenario, if with lower accuracy depending on the amount of missing data.
+This means that CCA implicitly assumes MCAR, and we see in \cite{Schafer2002} that we get unbiased inference in this scenario, though with lower accuracy depending on the amount of missing data.
 However, as with the height example, deviation from MCAR can introduce bias.
-See Schafer and Graham, ...
-for examples of how inference is affected.
-(CITE REVIEW PAPER FROM TURID).
+See \cite{Schafer2002} for examples of how inference is affected.
 
-	\textbf{Mean imputation:}
-	Another common practice, this technique involves substituting the missing values with the mean of the observed values for that variable.
-As stated in Schafer and Graham [2002], this has the effect of preserving the mean of the distribution one wishes to draw inference about, but distorts other parameters such as variance and covariance.
-Covariance especially is distorted due to mean imputation taking only the varible under consideration as input.
-In addition to this the technique breaks down when confronted with categorical variables where the mean does not exist (and it is clear how a majority vote imputation might skew the data).
-%Add some more math from either schafer and graham or another source
+	\subsubsection{Mean imputation}
+	Another common practice involves substituting the missing values with the mean of the observed values for that variable.
+As stated in Schafer and Graham \cite{Schafer2002}, this has the effect of preserving the mean of the distribution one wishes to draw inference about, but distorts other parameters such as variance and covariance.
+Covariance especially is distorted due to mean imputation taking only the variable under consideration as input.
+In addition to this, the technique breaks down when confronted with categorical variables for which a mean has a less obvious meaning, and it is clear how a majority vote imputation might skew the data.
 
-	\textbf{Last Observation Carried Forward (LOCF):x}
-	This method is mostly used in longitudinal studies where individuals drop out, withdraw, or are lost over the course of the studies.
+	\subsubsection{Last Observation Carried Forward}
+	Last Observation Carried Forward (LOCF) is mostly used in longitudinal studies where individuals drop out, withdraw, or are lost over the course of the studies.
 LOCF involves imputing the missing values for an individual with the last observed value for that individual.
-Leaving aside questions of precision, LOCF produces biased estimators in all except very specific MNAR circumstances (Kenward and Molenberghs 2009 get article).
-With LOCF we are effectively assuming that for each individual that drops out their responses do not change at all for the remainder of the study, a highly unrealistic assumption.
+By replicating already existing values, LOCF has a dampening impact on the variance estimate.
+Additionally, LOCF produces biased estimators in all except very specific MNAR circumstances \cite{KenwardM.G.andMolenberghs2009}.
+With LOCF, we are effectively assuming that for each individual that drops out, their responses do not change at all for the remainder of the study, a highly unrealistic assumption.
 It can be concluded, therefore, that LOCF is the worst imputation method of those mentioned here, and all the other methods are preferable.
-According [Missing data] "LOCV is neither valid under general assumptions nor based on statistical principles, it is not a sensible method, and should not be used".
+We round off by quoting \cite[P.
+40]{Molenberghs2015} that\enquote{LOCV is neither valid under general assumptions nor based on statistical principles, it is not a sensible method, and should not be used}.
 
-	\textbf{Regression:}
+	\subsubsection{Regression}
 	Regression imputation functions by regressing the variable(s) with missing entries on the fully observed variables, or subsets that are fully observed.
-To illustrate this, let us consider the simplest case, with one fully observed variable $x$ linearly related to a variable with missing data $y$.
-Fitting a regression of $y$ on $x$, we have$$Y_{i} = \beta_{0} + \beta_{1}x_{i} + \epsilon \qquad i\in(0,...,n)$$ obtaining estimates of the regression parameters $\hat{\beta_{0}}$ and $\hat{\beta_{1}}$ allows us to impute $E(y_{i}) = \hat{\beta_{0}} + \hat{\beta_{1}}x_{i}$.
-This is the best of the ad hoc methods, as it gives us much more plausible imputations that can vary.
+To illustrate this, let us consider the simplest case, with one fully observed variable $X$ linearly related to a variable $y
+	Y$ with missing data.
+Fitting a regression of $X$ on $Y$, we have$$Y_{i} = \beta_{0} + \beta_{1}x_{i} + \epsilon, \qquad \epsilon \sim N(0,1) \qquad i\in(0,...,n)$$ Obtaining estimates of the regression parameters $\hat{\beta_{0}}$ and $\hat{\beta_{1}}$ allows us to impute $E(y_{i}) = \hat{\beta_{0}} + \hat{\beta_{1}}x_{i}$.
+A variation on this method is to use the regression to specify the parameters of a normal distribution $Y ~ N(X\beta, \sigma_{y})$ from which to draw our imputation values.
+This is the best of the ad hoc methods, as it gives us much more plausible imputations that vary more like the true data \cite[P.
+37]{Molenberghs2015}.
 However, even in this simple case it yields imputations that tend to be less variable than the observed data.
-Therefore, using them in inference will most often lead to underestimated standard errors and p-values.
-If there are non-linear relationships in the data, this method's performance breaks down with regard to other aspects of inference as well.
+Therefore, using these imputations in inference will most often lead to underestimating standard errors and p-values.
+If there are non-linear relationships in the data, this method's performance breaks down.
 
-	For more in depth descriptions and examples of these methods we refer the reader to [missing data handbook].
-
-	\subsection{Statistically principled imputation}
-	By statistically principled, we mean motivated by an ability to preserves the statistical qualities in the data which we care about.
+	\section{Statistically principled imputation}
+	The methods in this section are motivated from a desire to preserves the distributional qualities in the data.
 For inference, this would be parameter values of a distribution and their variance and covariance.
 
-	\textbf{Single imputation}
-	Single imputation is the building block of MI, and ..
-Single imputation can be done in a variety of ways, but the most straightforward and easy to understand is the following: create a joint posterior distribution ...
-Draw from this distribution to fill in your dataset, and then do your analysis.
+	\subsection{Single imputation using conditional distributions}
+	Single imputation is the building block of MI, and many of its deficiencies are corrected for when expanded to MI.
+It can be done in a variety of ways, but the most straightforward and easy to understand is the following: create a joint posterior distribution of your missing data given your observed data, i.e.
+$f(Y^{m}|Y^{o}, M)$.
+Draw from this distribution to fill in your dataset, then perform your desired analysis.
+How you specify the missing data model can determine the efficacy of this method, but it is otherwise uncomplicated.
+We will therefore focus the majority of our analysis on the MI case.
 
-	\section{Multiple imputation}
+	\subsection{Multiple imputation}
 	One sizeable weakness of the earlier imputation methods is that they use points as estimates of distributions.
 Hence, subsequent measures of precision are biased downwards.
-Multiple imputation solves this problem by creating multiple imputed datasets drawn from a posterior distribution and then combines them in the final step.
-This preserves the distributional variation in the missing data, and ...
-Increases in available computation power has made MI a much more practical method to implement, as one of the weaknesses that earlier impeded its adoption was that it was computationally expensive.
+Multiple imputation solves this problem by creating multiple imputed datasets drawn from $f(Y^{m}|Y^{o}, M)$, running the desired analysis procedure on each, and then combining the results.
+This preserves the distributional variation in the missing data, and leads to better variance estimates.
+It can be quite computationally expensive, especially for large data sets, but recent increases in available computation power has made this concern less of an issue.
+Many previous objections to MI regarding inference concerned MI using a low number of imputed datasets $M$, typically between 5 and 10 \cite{Meng1994}.
+With larger $M$ and careful combining of the resulting analyses, one can achieve strong results in a large number of situations, see \cite{Xie2017}), \cite{Meng1994}, \cite{rubin_multiple_1996}, and \cite{madley-dowd_proportion_2019}.
 
 	Multiple imputation can be thought of as three distinct steps:
-	1) M imputed datasets are generated through repeated random draws from the predictive distribution of the missing values under a particular model for missingness.
-	2) The intended analysis is performed separately on all M datasets
-	3) The results of the analysis are combined using Rubin's or Meng's rules to create a single MI result.
-
-	Between imputation variance and within imputation variance
-
-	It should be noted that even though there is no strict mathematical proof that MI yields valid inference, the empirical evidence presents a strong case that MI is a useful tool.
+	\begin{enumerate}
+		\item $M$ imputed datasets are generated through repeated random draws from the predictive distribution of the missing values, $f(Y^{m}|Y^{o}, M)$, under a particular model for missingness.
+		\item The intended analysis is performed separately on all $M$ datasets
+		\item The results of the analyses are combined using Rubin's or Meng's rules (which we will describe shortly) to create a single MI result.
+	\end{enumerate}
+	It should be noted that even though there is no strict mathematical proof that MI yields valid inference in the case of finite $M$, the empirical evidence in the literature so far referenced presents a strong case that MI is a useful tool.
 Comparing it to the other methods available, one finds that alternatives are on the whole less effective.
 
-	\subsection{Specifying the imputation model}
+	\subsubsection{Specifying the imputation model}
 
-	We often want to include as many covariates as possible.
-X have shown that even mildly correlated auxiliary variables can improve subsequent inference [cite annotadet article].
-However, this can cause the imputation and and analyses models to be uncongenial.
+	To create an accurate model for the missing data, we often want to include as many covariates as possible.
+Madley-Dowd et al.
+[2019] have shown that even mildly correlated auxiliary variables can improve subsequent inference.
+By auxiliary, variables we mean variables not included in the analysis model.
+For example, imagine we have data on 1000 people's age, height, and weight.
+Suppose we wish to find the relationship between age and weight, but we are missing some of the weight data.
+Since height is correlated with weight, i.e.
+it contains some information about the distribution of weight, it makes sense to include it in our imputation model, despite not using it in our analysis.
+This leads us to make repeated random draws from $f(Y^{m}|Y^{o}, M, A)$, where $A$ represents auxiliary variables.
+Whether or not to include auxialiary variables can be either a cause and a cure for uncongeniality depending on the situation.
+Meng [1994] cautions against the use of less general imputation models (e.g.
+ignoring covariates/auxiliary variables).
 
-	\subsection{Rubin's rules and Meng's adaptations}
+	\subsubsection{Rubin's rules}
 
-	At step 2) of MI, we have M sets of parameter estimates and their variances $\hat{\theta}_{i},\hat{Var_{i}} \qquad i=0,...,M$.
+	At step 2 of MI, we have $M$ sets of parameter estimates and their variances: $\hat{\theta}_{i},\hat{Var_{i}}, \quad i=0,...,M$.
 The next step is to combine these into a single estimate.
-Under congeniality Rubin's combining rules suffice for giving us statistically valid estimates.
-The parameter estimate is simply the average $$\bar{\theta} = \frac{1}{M}\sum_{i=0}^{M}\hat{\theta}_{i}$$ and the variance associated with $\bar{\theta}$ is given by $$T = \bar{Var}+(1+\frac{1}{M}B)$$ where $$\bar{Var} = \frac{1}{M}\sum_{i=0}^{M}\hat{Var}_{i}$$ measures the within imputation variability and $$B = \frac{1}{M-1}\sum_{i=0}^{M}(\hat{\theta_{i}}-\bar{\theta})(\hat{\theta_{i}}-\bar{\theta})^{T}$$ measures the between imputation variability.
-The adjustment $(1+\frac{1}{M})$ is due to the finite number of imputations.
-If we let $M\rightarrow\infty$ then...
+Under congeniality, Rubin's combining rules suffice for giving us statistically valid estimates \cite{Meng1994}.
+The parameter estimate is simply the average $$\bar{\theta} = \frac{1}{M}\sum_{i=0}^{M}\hat{\theta}_{i}$$ and the variance associated with $\bar{\theta}$ is given by
 
-	If models are uncongenial, we use Meng's rules.
+	\begin{equation}
+		T = \overline{Var}+(1+\frac{1}{M}B)\tag{*}
+	\end{equation}
 
-	\subsection{Potential pitfalls and criticism}
+	where $$\overline{Var} = \frac{1}{M}\sum_{i=0}^{M}\hat{Var}_{i}$$ measures the within imputation variability and $$B = \frac{1}{M-1}\sum_{i=0}^{M}(\hat{\theta_{i}}-\bar{\theta})(\hat{\theta_{i}}-\bar{\theta})^{T}$$ measures the between imputation variability.
+The adjustment $(1+\frac{1}{M}B)$ in equation $(*)$ is due to the finite number of imputations.
+If we let $M\rightarrow\infty$ then we achieve asymptotic equality to the true variance estimate.
+
+	Xia and Meng [2017] state that \enquote{When the imputation model class and the (embedded) analyst’s model class
+		differ, the behavior of Rubin’s rules becomes very complicated, capable of producing inconsistent variance estimators, a matter that has received recurrent
+		criticisms}.
+They argue that in the uncongenial case, which is often the rule rather than the exception, doubling Rubin's variance measure will never underestimate the true variance.
+This holds for any degree of uncongeniality, and provides a surprisingly simple solution to the problem of uncongeniality.
+For univariate data, Xia and Meng  [2017] suggest replacing $T$ with $\tilde{T}= (\sqrt{\overline{Var}}+\sqrt{B})^{2}+\frac{1}{M}B$.
+These methods may be viewed as highly conservative, but in the context of MI where uncongeniality between analysts and imputer is unknown and potentially large, it pays to be as insured as possible against invalid inference.
+However, even if these methods are not used and the models are uncongenial, we note that Meng [1994] has shown that MI is still superior to CCA.
+
+	\subsubsection{Potential pitfalls and criticism}
 	The following arguments apply for imputation in general, but we will consider them in the context of MI.
-When there is missing data in several variables of different types ,for example continuous and binary, and the data is structured, specifying the imputation model can be quite difficult.
-This is because it has to be a joint model...
+When there is missing data in several variables of different types, for example continuous and binary, and there exists hierarchical structure in the data, specifying the imputation model can be quite difficult.
+A solution to the problem is using chained equations, where one has a univariate model for each variable with missing data and iteratively draws from this.
+Despite good empirical results, this method lacks the theoretical justification of drawing from a fully specified conditional distribution \cite{Little2019}, and so is at best a partial solution.
+Another danger is that improper specification of the imputation model can cause uncongeniality and so one needs to exercise care when using MI.
+Finally, imputation procedures based on drawing from conditional distributions tend to assume that these distributions are normal.
+In certain cases this requires us to transform the data.
 
-	. Among all the
-	criticisms of multiple imputation, the recent work
-	of Fay (1991, 1992) is the most intense, as it is
-	directed at the validity of multiple-imputation in-
-	ference in practice. Fay provided several examples
-	(see also Kott, 1992) to show that the variance esti-
-	mator obtained from the repeated-imputation combining rules disagrees asymptotically with the sam-
-	pling variance of the repeated-imputation estimator,
-	even when the imputation model is correctly speci-
-	fied. [MENG]
-
-	%^fixed with congeniality
-
-	Even Rubin (1996) admits that MI
-
-	Taking all this into account, it is important to see MI in context of the options available to someone working with data.
-As we have established earlier, simply ignoring missing data can lead  to serious biases, while alternative methods yield similar if not inferior performance in certain cases.
-As long as a potential user is aware of the potential pitfalls, MI is the superior method for dealing with missing data.
-
-	\subsection{When to use multiple imputation?}
-	Computational resources and time to create the full joint imputation model. (Handbook 236)
-
-	A common measure used to decide when to use multiple imputation is percentage of missing data.
-
-	However, Madley-Dowd et. al. (Proportion of missing data paper) propose a different measure.
-
-	MI can be used to perform sensitivity analysis by varying the imputation model.
+	Taking all this into account, it is important to see MI in the context of the options available to someone working with data.
+As we have established earlier, simply ignoring missing data can lead  to serious biases, while alternative methods tend to negatively effect estimates and/or their variance.
+Since there is no option to \enquote{do nothing}, MI presents itself as a useful way of dealing with missing data.
 
 	\section{Experiments}
-	In this section we will use Madley-Dowd et. al.'s missingness measure to...
 
 	\subsection{MICE}
-	''The
-	increased flexibility in modeling these conditional distributions may outweigh
-	the lack of clear theoretical justification of the method.'' Rubin
-
-	Lee and Carlin  [2010] argue that MICE and MVN have comparable performance.
-
-	We have chosen to use MICE for the experiments in this paper for two main reasons.
-One is the aformentioned performance and flexibility, while the other is that MICE is unlicenced and open source in contrast to other technqieus like drawing from a fully specified joint distribution found in statistical packages like STATA.
-In the interest of democratizing MI while retaining the desired performance, MICE was found to be superior.
+	We have chosen to use the MICE package for R \cite{Buuren2011} for our experiments.
+There are two main reasons for this.
+One is the performance: MICE is able to deal with the complex scenario of mixed variable types as described earlier.
+MICE functions by drawing from univariate conditional distributions for all the variables with missing data, i.e for a variable $Y_{i}$ we draw from the predictive distribution $p(y_{i}|y_{0}, ...
+y_{j}), \quad i\neq j$ for each i in an iterative process.
+This solves the problem of complex data mentioned earlier, and despite the lack of theoretical justification, its performance is comparable to the fully specified conditional distribution approach \cite{Lee}, \cite{Little2019}.
 
@@
 
-	%Remember to use Rubin's definitons of statistical validity in evaluating the methods.
+	In this section we will provide a simple experiment to explore the efficacy of MI.
+We will look at a simple linear regression with a single covariate defined by the model $Y_{i} = \beta_{0} + \beta_{1}x_{i} + \epsilon, \qquad \epsilon \sim N(0,1) \qquad i\in(0,...,n)$.
+Missingness will be simulated in both $X$ and $Y$.
+Data was generated by drawing from a bivariate normal distribution $$[Y X] \sim N(\begin{bmatrix}
+	0 \\
+	0 \\
+	\end{bmatrix},\begin{bmatrix}
+	1 & 0.8 \\
+	0.8 & 1 \\
+	\end{bmatrix})$$ The true regression coefficient is then $\beta_{1} = 0.8$.
+In our experiment we draw 1000 data points from this distribution.
+MCAR was simulated by removing the first p values in the data set.
+MAR was simulated using the logit function where the probability of an observation being missing was equal to $\frac{1}{1+e^{-(Y + \alpha)}}$ when simulating missingness in X, and $\frac{1}{1+e^{-(X + \alpha)}}$ when simulating missingness in $Y$.
+$\alpha$ was chosen such that we achieved the desired percentage of missing data.
+MNAR was simulated using the same logit function, but in this case the probability of $Y$ missing was dependent on $Y$.
+The number of imputations $M$ was set to 20.
+We will look at the following metrics to evaluate the performance of MI:
 
-	"The MCAR missingness mechanism
-	removed the first p observations such that pn gives the
-	required proportion of missing data. MAR missingness
-	was simulated under a logistic regression model using
+	\begin{enumerate}
+		\item Size of the confidence interval of $\beta_{1}$
+		\item Coverage of the confidence interval
+		\item Bias
+	\end{enumerate}
 
-	The value of a was manipulated for the different simu-
-	lation settings to provide the required proportion of missing
-	data on average across data sets." from (P. Madley-Dowd et a)
+	We will compare the performance of MI to an equivalent analysis performed with CCA for reference. Our experiment was then conducted in the following manner:
 
-	Also check out their motivation for the generating of the dataa
+	\begin{enumerate}
+		\item Draw 1000 observations from the distributions of $[Y X]$
+		\item Simulate 5, 10, 20, 40, 60, 80, and 90 percent missingness under different missingness mechanisms.
+		\item Create linear models for each of these 7 datasets based on CCA and MI
+		\item Repeat 1000 times
+		\item Evaluate the results based on the above metrics.
+	\end{enumerate}
 
-	In the following examples we will compare our models' performance across different data types.
-These are continuous, categorical (binary and multicategory), time to event data, and timeseries.
-The latter two
+	\subsubsection{MCAR}
 
-	\subsubsection{Inference}
-	Test statistics
-	Different tests
-	x
-	Congenial vs uncongenial models, see the annotated paper for how they did things with regard to imputation vs analysis.
+	\begin{figure}[H]
+		\centering
+		\subfloat[$\beta_{1}$ estimate with CCA]{\includegraphics[width=0.4\textwidth]{mcar_estimate_cca.png}}
+		\quad
+		\subfloat[$\beta_{1}$ estimate with MI]{\includegraphics[width=0.4\textwidth]{mcar_estimate_mi.png}}
+		\\
+		\subfloat[Coverage]{\includegraphics[width=0.3\textwidth]{mcar_coverage.png}}
+		\quad
+		\subfloat[Average CI size]{\includegraphics[width=0.3\textwidth]{mcar_ci.png}}
+		\quad
+		\subfloat[Average bias]{\includegraphics[width=0.3\textwidth]{mcar_bias.png}}
+		\caption{Average results for MCAR missingness in $Y$ over 1000 repetitions of the analysis procedure described in 6.2.}
+	\end{figure}
 
-	\subsubsection{Prediction}
-	With large publicly available datasets on the one hand, and laws like GDPR requiring deleting certain kinds of data on the other, multiple imputation seems like a strong candidate for improving the predictive quality of models.
-Most of the literature focuses on inference, so there is comparatively little comparing the performance of different predictive models.
-Here we will compare three basic model architectures, multiple linear regression, boosted trees, and a simple neural network, to see how they perform with and without multiple imputation.
+	Looking first at the $\beta_{1}$ estimates and their confidence intervals, we see that there is little difference between the methods other than the confidence interval of CCA growing at a much faster pace than that of MI as missingness increases.
+Next we observe that Bias is lower for MI at low missingness, while higher at high missingness.
+Additionally, coverage decreases substantially for MI as missingness increases.
 
-	\subsubsection{Classification}
-	Like with the prediction case above, classification is an area in which the litterature on missing data is sparse and the same conditions apply.
-In this context we will compare logistic regression, boosted trees, and neural networks.
+	Overall it seems like MI provides some benefit over CCA in a low missingness scenario, while CCA seems superior otherwise.
+The poor performance of MI at higher amounts of missingness suggests an instability in the implementation, whether in MICE or the supplementary code.
+Another possibility is that at 90 percent missingness our estimates for the conditional distribution of $Y$ become too inaccurate.
 
-	\subsection{Sensitivity analysis and uncongenial models}
+	\subsubsection{MAR}
 
-	\subsection{Real data}
+	\begin{figure}[H]
+		\centering
+		\subfloat[$\beta_{1}$ estimate with CCA]{\includegraphics[width=0.4\textwidth]{mar_reg_x_estimate_cca.png}}
+		\quad
+		\subfloat[$\beta_{1}$ estimate with MI]{\includegraphics[width=0.4\textwidth]{mar_reg_x_estimate_mi.png}}
+		\\
+		\subfloat[Coverage]{\includegraphics[width=0.3\textwidth]{mar_reg_x_coverage.png}}
+		\quad
+		\subfloat[Average CI size]{\includegraphics[width=0.3\textwidth]{mar_reg_x_ci.png}}
+		\quad
+		\subfloat[Average bias]{\includegraphics[width=0.3\textwidth]{mar_reg_x_bias.png}}
+		\caption{Average results for MAR missingness in $X$ over 1000 repetitions of the analysis procedure described in 6.2.}
+	\end{figure}
+	For missingness in $X$ we find that MI provides substantial benefits in reducing bias and increasing coverage while at the same time reducing the size of confidence intervals.
+This is in line with earlier theory and results described in literature.
 
-	Description of the data
+	\begin{figure}[H]
+		\centering
+		\subfloat[$\beta_{1}$ estimate with CCA]{\includegraphics[width=0.4\textwidth]{final_estimate_cca.png}}
+		\quad
+		\subfloat[$\beta_{1}$ estimate with MI]{\includegraphics[width=0.4\textwidth]{final_estimate_mi.png}}
+		\\
+		\subfloat[Coverage]{\includegraphics[width=0.3\textwidth]{final_coverage.png}}
+		\quad
+		\subfloat[Average CI size]{\includegraphics[width=0.3\textwidth]{final_ci.png}}
+		\quad
+		\subfloat[Average bias]{\includegraphics[width=0.3\textwidth]{final_bias.png}}
+		\caption{Average results for MAR missingness in $Y$ over 1000 repetitions of the analysis procedure described in 6.2.}
+	\end{figure}
 
-	HUNT Norwegian data
+	In contrast to missingness in $X$, we do not find any benefits from using MI when missingness is in $Y$.
+In fact, we observe that CCA outperforms MI on all of our metrics.
+The strong performance of CCA on its own is surprising, and suggests that the data is MCAR rather than MAR despite use of the logit function to remove data.
 
-	Andreas data webSite
-	balanced number of classes for classification
+	\subsubsection{MNAR}
 
-	\subsection{TODO}
-	Difference between baysian and frequentist paradigm.
-	Citation example:
-	\cite[P.10]{grund_pooling_2016}.
+	\begin{figure}[H]
+		\centering
+		\subfloat[$\beta_{1}$ estimate with CCA]{\includegraphics[width=0.4\textwidth]{mnar_final_estimate_cca.png}}
+		\quad
+		\subfloat[$\beta_{1}$ estimate with MI]{\includegraphics[width=0.4\textwidth]{mnar_final_estimate_mi.png}}
+		\\
+		\subfloat[Coverage]{\includegraphics[width=0.3\textwidth]{mnar_final_coverage.png}}
+		\quad
+		\subfloat[Average CI size]{\includegraphics[width=0.3\textwidth]{mnar_final_ci.png}}
+		\quad
+		\subfloat[Average bias]{\includegraphics[width=0.3\textwidth]{mnar_final_bias.png}}
+		\caption{Average results for MNAR missingness in $Y$ over 1000 repetitions of the analysis procedure described in 6.2.}
+	\end{figure}
+	We observe that both methods have a similar performance until they reach 60 percent missingness where their performance drastically diverges.
+Under MNAR, without specifying the missingness mechanism in the imputation model there is a high chance that it will be uncongenial with the analysis model.
+If this is the case, then this would go some way in explaining the poor performance of MI at hgih missingness.
 
-	\bibliography{bib2}	
+	\subsection{Should we use multiple imputation?}
+	Overall MI has shown quite poor performance in our experiments.
+Theory suggests that CCA should at best equal the performance of MI, and only in the MCAR scenario.
+We see from our experiments that multiple imputation loses some of its efficacy at extremely high levels of missingness.
+However, the research of Madley-Dowd et.
+al.
+[2019] suggest otherwise, showing that MI can yield large benefits for bias and standard error even at high missingness levels.
+Madley-Dowd et.
+al.
+make use of auxiliary covariates, and so it is plausible that this has an impact on the performance of MI relative to our experiment.
+Given the vast amount of theory and experimental results on this topic, some of which has been cited in this thesis, the likely scenario is that the implementation has skewed the results.
+
+	\section{Conclusion and further work}
+	This thesis has shown that missing data is an important consideration in data analysis by the effect it has on inference.
+The mechanism behind missing data is key to addressing it, and this thesis has characterised three possible such mechanisms.
+Imputation has been established as a statistically valid technique for addressing missing data.
+Various common and intuitive imputation methods such as Complete Case Analysis and Mean Imputation have been evaluated, with Multiple Imputation found to be the superior method.
+Theory and literature suggest strong benefits from using MI, however, this was not fully borne out in our own experiments.
+Nonetheless, the substantial amount of literature in favour of MI suggests that our experimental results were due to an error in implementation rather than an error with the method.
+Further study will be needed to verify this.
+
+	Possible future investigations could be made into the effect of MI on prediction and classification, especially in the context of machine learning models.
+Another interesting avenue of exploration is replacing the conditional distributions with other modeling tools such as trees or neural networks.
+Perhaps the messy nature of much of modern data will make the underlying assumptions of conditional distributions untenable.
+
+	\bibliography{final_5_bib}	
 	\bibliographystyle{apalike}

### 2020-06-11 experiments.Rmd
Finished

--- 2020-02-25 experiments.Rmd
@@
 ---
-```{r}
-
-```
+TODOS:
+Make simulation with cross correlated variables
+Incorporate FMI
+Add classification and predictionY
 
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = FALSE)
 ```
@@
 library(mvtnorm)
+library(MASS)
+library(tidyverse) 
+library(ggplot2)
+```
+```{r}
+'?'(mvtnorm)
+mu = c(1,1)
+sigma = matrix(c(c(1, 0.6), c(0.6, 1)), nrow = 2, ncol = 2)
+data = rmvnorm(n, mu, sigma)
+data = data.frame("y_1" = data[,1], "x_1" = data[,2])
+reference = lm(y_1 ~x_1, data = data)
+summary(reference)
+
+data_1_sixty = data
+sixty_percent = 1/(1+exp(-(data$x_1-1.5)))
+mean(sixty_percent)
+for (i in 1:length(sixty_percent)){
+  val = rbinom(1, 1, sixty_percent[i])
+  #print(val)
+  if(val == 0 ){
+    data_1_sixty$y_1[i] = NA
+  }
+}
+hist(data_1_sixty$y_1)
+hist(data$y_1)
+colMeans(data)
+colMeans(na.omit(data_1_sixty))
+mean(data_1_sixty$x_1)
+data_1_sixty
 ```
@@
 ```{r}
-#Setting vals for multivariate continous vars
-mu = 8
-sigma = 2
-help(mvtnorm)
+#Setting number of entries
+n = 500
+iters = 20
+
 ```
@@
 ```{r}
-#Generating binary data
-n = 1000
-binary_categorical_vector = round(runif(n, min = 0, max = 1))
+n = 500
+x_1 = rnorm(n, 1, 1)
+b_1 = 8
+xb = x_1*b_1+b_0
+sigma_i = 10
+y_1 = rnorm(n, xb, sigma_i)
+df_1_full = data.frame(y_1, x_1)
+reference = lm(y_1 ~x_1, data = df_1_full)
+summary(reference)
+
 ```
+
+```{r}
+#Removing vals
+five_percent = 1/(1+exp(-(df_1_full$x_1 +2.5)))
+mean(five_percent)
+ten_percent = 1/(1+exp(-(df_1_full$x_1 + 1.6)))
+mean(ten_percent)
+twenty_percent = 1/(1+exp(-(df_1_full$x_1+0.7)))
+mean(twenty_percent)
+forty_percent = 1/(1+exp(-(df_1_full$x_1-0.5)))
+mean(forty_percent)
+sixty_percent = 1/(1+exp(-(df_1_full$x_1-1.5)))
+mean(sixty_percent)
+eighty_percent = 1/(1+exp(-(df_1_full$x_1-2.65)))
+mean(eighty_percent)
+ninety_percent = 1/(1+exp(-(df_1_full$x_1 -3.5)))
+mean(ninety_percent)
+vals = 0
+for(i in ninety_percent){
+  if(i<0.1){
+    vals = vals+1
+  }
+}
+
+df_1_five = df_1_full
+for (i in 1:length(five_percent)){
+  val = rbinom(1, 1, five_percent[i])
+  #print(val)
+  if(val == 0 ){
+    df_1_five$y_1[i] = NA
+  }
+}
+
+df_1_ten = df_1_full
+for (i in 1:length(ten_percent)){
+  val = rbinom(1, 1, ten_percent[i])
+  #print(val)
+  if(val == 0 ){
+    df_1_ten$y_1[i] = NA
+  }
+}
+
+df_1_twenty = df_1_full
+for (i in 1:length(twenty_percent)){
+  val = rbinom(1, 1, twenty_percent[i])
+  #print(val)
+  if(val == 0 ){
+    df_1_twenty$y_1[i] = NA
+  }
+}
+
+df_1_forty = df_1_full
+for (i in 1:length(forty_percent)){
+  val = rbinom(1, 1, forty_percent[i])
+  #print(val)
+  if(val == 0 ){
+    df_1_forty$y_1[i] = NA
+  }
+}
+
+df_1_sixty = df_1_full 
+for (i in 1:length(sixty_percent)){
+  val = rbinom(1, 1, sixty_percent[i])
+  #print(val)
+  if(val == 0 ){
+    df_1_sixty$y_1[i] = NA
+  }
+}
+
+df_1_eighty = df_1_full
+for (i in 1:length(eighty_percent)){
+  val = rbinom(1, 1, eighty_percent[i])
+  #print(val)
+  if(val == 0 ){
+    df_1_eighty$y_1[i] = NA
+  }
+}
+
+df_1_ninety = df_1_full
+val = rbinom(500,1, ninety_percent)
+for (i in 1:length(val)){
+  if(val[i] == 0 ){
+    df_1_ninety$y_1[i] = NA
+  }
+}
+hist(df_1_ninety$y_1)
+
+df_1_ninety = df_1_full
+for (i in 1:length(eighty_percent)){
+  set.seed(i)
+  val = rbinom(1, 1, ninety_percent[i])
+  #print(val)
+  if(val == 0 ){
+    df_1_ninety$y_1[i] = NA
+  }
+}
+hist(df_1_ninety$y_1)
+```
+```{r}
+df_1_five = df_1_full
+val = rbinom(500,1, five_percent)
+for (i in 1:length(val)){
+  if(val[i] == 0 ){
+    df_1_five$y_1[i] = NA
+  }
+}
+
+df_1_ten = df_1_full
+val = rbinom(500,1, ten_percent)
+for (i in 1:length(val)){
+  if(val[i] == 0 ){
+    df_1_ten$y_1[i] = NA
+  }
+}
+
+df_1_twenty = df_1_full
+val = rbinom(500,1, twenty_percent)
+for (i in 1:length(val)){
+  if(val[i] == 0 ){
+    df_1_twenty$y_1[i] = NA
+  }
+}
+
+ df_1_forty = df_1_full
+val = rbinom(500,1, forty_percent)
+for (i in 1:length(val)){
+  if(val[i] == 0 ){
+    df_1_forty$y_1[i] = NA
+  }
+}
+
+  df_1_sixty = df_1_full
+val = rbinom(500,1, sixty_percent)
+for (i in 1:length(val)){
+  if(val[i] == 0 ){
+    df_1_sixty$y_1[i] = NA
+  }
+}
+
+  df_1_eighty = df_1_full
+val = rbinom(500,1, eighty_percent)
+for (i in 1:length(val)){
+  if(val[i] == 0 ){
+    df_1_eighty$y_1[i] = NA
+  }
+}
+
+  df_1_ninety = df_1_full
+val = rbinom(500,1, ninety_percent)
+for (i in 1:length(val)){
+  if(val[i] == 0 ){
+    df_1_ninety$y_1[i] = NA
+  }
+}
+'?'(dnorm)
+pnorm(df_1_full$x_1, mean = 1, sd = 1)
+
+```
+
+```{r}
+set.seed(1)
+
+  val =pnorm(df_1_full$x_1, mean = 1, sd = 1)
+df_1_five = df_1_full
+
+for (i in 1:length(val)){
+  if(val[i] < 0.5 ){
+    df_1_five$y_1[i] = NA
+  }
+}
+
+df_1_ten = df_1_full
+
+for (i in 1:length(val)){
+  if(val[i] < 0.1 ){
+    df_1_ten$y_1[i] = NA
+  }
+}
+
+df_1_twenty = df_1_full
+
+for (i in 1:length(val)){
+  if(val[i] < 0.2 ){
+    df_1_twenty$y_1[i] = NA
+  }
+}
+
+ df_1_forty = df_1_full
+
+for (i in 1:length(val)){
+  if(val[i] < 0.4 ){
+    df_1_forty$y_1[i] = NA
+  }
+}
+
+  df_1_sixty = df_1_full
+
+for (i in 1:length(val)){
+  if(val[i] < 0.6 ){
+
+    df_1_sixty$y_1[i] = NA
+  }
+}
+   vals = 0
+  df_1_eighty = df_1_full
+for (i in 1:length(val)){
+  if(val[i] < 0.8 ){
+    vals= vals+1
+    df_1_eighty$y_1[i] = NA
+  }
+}
+
+  df_1_ninety = df_1_full
+for (i in 1:length(val)){
+  if(val[i] < 0.9 ){
+    df_1_ninety$y_1[i] = NA
+  }
+}
+vals
+hist(df_1_full$y_1)
+hist(df_1_sixty$y_1)
+hist(df_1_five$y_1)
+hist(df_1_ten$y_1)
+hist(df_1_twenty$y_1)
+hist(df_1_forty$y_1)
+
+hist(df_1_eighty$y_1)
+hist(df_1_ninety$y_1)
+
+```
+
+```{r echo=FALSE}
+#Generating data according to Turid's recommended method
+#One cont normal variables
+n = 500
+iters = 15
+imputations = 100
+x_1 = rnorm(n, 1, 1)
+b_1 = 1
+b_0 = 0
+xb = x_1*b_1 +b_0
+sigma_i = 10
+y_1 = rnorm(n, xb, sigma_i)
+df_1_full = data.frame(y_1, x_1)
+#Removing vals
+five_percent = 1/(1+exp(-(df_1_full$x_1 +2.5)))
+mean(five_percent)
+ten_percent = 1/(1+exp(-(df_1_full$x_1 + 1.6)))
+mean(ten_percent)
+twenty_percent = 1/(1+exp(-(df_1_full$x_1+0.7)))
+mean(twenty_percent)
+forty_percent = 1/(1+exp(-(df_1_full$x_1-0.5)))
+mean(forty_percent)
+sixty_percent = 1/(1+exp(-(df_1_full$x_1-1.5)))
+mean(sixty_percent)
+eighty_percent = 1/(1+exp(-(df_1_full$x_1-2.65)))
+mean(eighty_percent)
+ninety_percent = 1/(1+exp(-(df_1_full$x_1 -3.5)))
+mean(ninety_percent)
+
+removers = c(five_percent, ten_percent, twenty_percent, 
+             forty_percent, sixty_percent, eighty_percent, ninety_percent)
+
+df_1_five = df_1_full
+for (i in 1:length(five_percent)){
+  val = rbinom(1, 1, five_percent[i])
+  #print(val)
+  if(val == 0 ){
+    df_1_five$y_1[i] = NA
+  }
+}
+
+df_1_ten = df_1_full
+for (i in 1:length(ten_percent)){
+  val = rbinom(1, 1, ten_percent[i])
+  #print(val)
+  if(val == 0 ){
+    df_1_ten$y_1[i] = NA
+  }
+}
+
+df_1_twenty = df_1_full
+for (i in 1:length(twenty_percent)){
+  val = rbinom(1, 1, twenty_percent[i])
+  #print(val)
+  if(val == 0 ){
+    df_1_twenty$y_1[i] = NA
+  }
+}
+
+df_1_forty = df_1_full
+for (i in 1:length(forty_percent)){
+  val = rbinom(1, 1, forty_percent[i])
+  #print(val)
+  if(val == 0 ){
+    df_1_forty$y_1[i] = NA
+  }
+}
+
+df_1_sixty = df_1_full 
+for (i in 1:length(sixty_percent)){
+  val = rbinom(1, 1, sixty_percent[i])
+  #print(val)
+  if(val == 0 ){
+    df_1_sixty$y_1[i] = NA
+  }
+}
+
+df_1_eighty = df_1_full
+for (i in 1:length(eighty_percent)){
+  val = rbinom(1, 1, eighty_percent[i])
+  #print(val)
+  if(val == 0 ){
+    df_1_eighty$y_1[i] = NA
+  }
+}
+
+df_1_ninety = df_1_full
+for (i in 1:length(ninety_percent)){
+  val = rbinom(1, 1, ninety_percent[i])
+  #print(val)
+  if(val == 0 ){
+    df_1_ninety$y_1[i] = NA
+  }
+}
+
+#CCA
+cca_1_5 = lm(y_1 ~x_1, data = df_1_five)
+cca_1_10 = lm(y_1 ~ x_1, data = df_1_ten )
+cca_1_20 = lm(y_1 ~ x_1, data = df_1_twenty)
+cca_1_40 = lm(y_1 ~ x_1, data = df_1_forty)
+cca_1_60 = lm(y_1 ~ x_1, data = df_1_sixty)
+cca_1_80 = lm(y_1 ~ x_1, data = df_1_eighty)
+cca_1_90 = lm(y_1 ~ x_1, data = df_1_ninety)
+
+'?'(mice)
+imp_1_5 = mice(df_1_five, m = imputations, maxit = iters, printFlag = FALSE)
+imp_1_10 = mice(df_1_ten, m = imputations,maxit = iters, printFlag = FALSE)
+imp_1_20 = mice(df_1_twenty,m = imputations, maxit = iters, printFlag = FALSE)
+imp_1_40 = mice(df_1_forty, m = imputations,maxit = iters, printFlag = FALSE)
+imp_1_60 = mice(df_1_sixty, m = imputations,maxit = iters, printFlag = FALSE)
+imp_1_80 = mice(df_1_eighty,m = imputations, maxit = iters, printFlag = FALSE)
+imp_1_90 = mice(df_1_ninety, m = imputations,maxit = iters, printFlag = FALSE)
+
+fit_1_5 = with(imp_1_5, lm(y_1 ~x_1))
+fit_1_10 = with(imp_1_10, lm(y_1 ~x_1))
+fit_1_20 = with(imp_1_20, lm(y_1 ~x_1))
+fit_1_40 = with(imp_1_40, lm(y_1 ~x_1))
+fit_1_60 = with(imp_1_60, lm(y_1 ~x_1))
+fit_1_80 = with(imp_1_80, lm(y_1 ~x_1))
+fit_1_90 = with(imp_1_90, lm(y_1 ~x_1))
+
+pooled_1_5 = pool(fit_1_5)
+pooled_1_10 = pool(fit_1_10)
+pooled_1_20 = pool(fit_1_20)
+pooled_1_40 = pool(fit_1_40)
+pooled_1_60 = pool(fit_1_60)
+pooled_1_80 = pool(fit_1_80)
+pooled_1_90 = pool(fit_1_90)
+
+r_sq_1_5 = pool.r.squared(fit_1_5)
+r_sq_1_10 = pool.r.squared(fit_1_10)
+r_sq_1_20 = pool.r.squared(fit_1_20)
+r_sq_1_40 = pool.r.squared(fit_1_40)
+r_sq_1_60 = pool.r.squared(fit_1_60)
+r_sq_1_80 = pool.r.squared(fit_1_80)
+r_sq_1_90 = pool.r.squared(fit_1_90)
+
+w = summary(pooled_1_5, conf.int=TRUE)
+w
+beta_ci(summary(pooled_1_5,  conf.int=TRUE))
+```
+
+```{r}
+#Old
+#Function for calculating confidence intervals for betas
+#beta_ci = function(model_summ, lower, upper){
+#  print(model_summ)
+#  std = model_summ$std.error
+#  estim = model_summ$estimate
+#  ci_ = c()
+#  for(i in 1:length(estim)){
+#    print(c(estim[i] - 1.96*std[i], estim[i] + 1.96*std[i]))
+#    ci_ = rbind(ci_, c(estim[i] - 1.96*std[i], estim[i] + 1.96*std[i]))
+#  }
+#  #print("We are at the ifs")
+#  if(lower){
+#    return(ci_[1,])
+#  }
+#  if(upper){
+#    return(ci_[2,])
+#  }
+#  return(ci_)
+#}
+
+#New
+beta_ci = function(model_summ){
+  #print("We are in the function")
+  w = model_summ
+  ci_ = list("lower" = w$`2.5 %`[2], "upper" = w$`97.5 %`[2])
+  return(ci_)
+}
+```
+
+```{r}
+r_sq_cca = c(summary(cca_1_5)$r.squared, summary(cca_1_10)$r.squared , summary(cca_1_20)$r.squared, summary(cca_1_40)$r.squared, summary(cca_1_60)$r.squared, summary(cca_1_80)$r.squared, summary(cca_1_90)$r.squared)
+r_sq_total = c(r_sq_1_5[1], r_sq_1_10[1], r_sq_1_20[1], r_sq_1_40[1], r_sq_1_60[1], r_sq_1_80[1], r_sq_1_90[1])
+
+beta_1_mi =c(summary(pooled_1_5)$estimate[2], summary(pooled_1_10)$estimate[2], summary(pooled_1_20)$estimate[2], summary(pooled_1_40)$estimate[2], summary(pooled_1_60)$estimate[2], summary(pooled_1_80)$estimate[2], summary(pooled_1_90)$estimate[2])
+
+beta_1_mi_ci = rbind(beta_ci(summary(pooled_1_5,  conf.int=TRUE)), beta_ci(summary(pooled_1_10,  conf.int=TRUE)), beta_ci(summary(pooled_1_20,  conf.int=TRUE)), beta_ci(summary(pooled_1_40,  conf.int=TRUE)), beta_ci(summary(pooled_1_60,  conf.int=TRUE)), beta_ci(summary(pooled_1_80,  conf.int=TRUE)), beta_ci(summary(pooled_1_90,  conf.int=TRUE)))
+beta_1_mi_ci = data.frame("lower" = unlist(beta_1_mi_ci)[1:7], "upper" =unlist(beta_1_mi_ci)[8:14])
+
+beta_1_1_cca_ci = rbind(confint(cca_1_5)[2,], confint(cca_1_10)[2,], confint(cca_1_20)[2,], confint(cca_1_40)[2,], confint(cca_1_60)[2,], confint(cca_1_80)[2,], confint(cca_1_90)[2,])
+
+beta_both_1_cca = rbind(cca_1_5$coefficients, cca_1_10$coefficients, cca_1_20$coefficients, cca_1_40$coefficients, cca_1_60$coefficients, cca_1_80$coefficients, cca_1_90$coefficients)
+summary(cca_1_90)
+miss = c(5,10, 20, 40,60,80,90)
+
+df_1_beta = data.frame("Percent_missing" = miss, "beta_1_mi" = beta_1_1_mi, "beta_1_mi_lower" = beta_1_mi_ci$lower, "beta_1_mi_upper" = beta_1_mi_ci$upper,  "beta_0_cca" = beta_both_1_cca[,1], "beta_1_cca" = beta_both_1_cca[,2], "beta_1_cca_lower" =beta_1_1_cca_ci[,1], "beta_1_cca_upper" =beta_1_1_cca_ci[,2])
+#citation("mice")
+
+```
+
+```{r}
+#Plotting
+
+p<- ggplot(df_1_beta, aes(x=Percent_missing, y=beta_1_cca)) + 
+  geom_line() +
+  geom_point()+
+  geom_errorbar(aes(ymin=beta_1_cca_lower, ymax=beta_1_cca_upper), width=.2,
+                 position=position_dodge(0.05))
+t<- ggplot(df_1_beta, aes(x=Percent_missing, y=beta_1_mi)) + 
+  geom_line() +
+  geom_point()+
+  geom_errorbar(aes(ymin=beta_1_mi_lower, ymax=beta_1_mi_upper), width=.2,
+                 position=position_dodge(0.05))
+print(p)
+print(t)
+df_1_beta
+```
+
+```{r}
+results_5 =  data.frame("beta_0_mi" = double() , "beta_0_mi_lower" = double(), "beta_0_mi_upper" = double(), "beta_1_mi" = double(), "beta_1_mi_lower" = double(), "beta_1_mi_upper" = double(),  "beta_0_cca" = double(),  "beta_0_cca_lower"= double(), "beta_0_cca_upper" =double(), "beta_1_cca" = double(), "beta_1_cca_lower" =double(), "beta_1_cca_upper" =double())
+str(results_5)
+results_5 = rbind(c(1.0,1.0,0.0,0.0,0.0,0.0,0.0,0.0,0.0,0.0,0.0,0.0,0.0), results_5)
+results_5
+x = c(0,0)
+x = rbind(c(1,1), x)
+x = rbind(c(2,2), x)
+x = data.frame(x)
+
+```
+
+```{r}
+write.table(results_5, file='/home/marius/Downloads/Test/results_5.csv', row.names=FALSE)
+t = df_1_beta
+t$Percent_missing = NULL
+write.table(t[1,], file = "/home/marius/Downloads/Test/results_5.csv", append = TRUE, quote = FALSE,
+    col.names = FALSE, row.names = FALSE)
+summary(cca_model)$coefficients[3,4]
+summary(pooled_1_20)[2,6]
+x_1 = rnorm(n, 10, 1)
+b_1 = 1
+b_0 = 0
+xb = x_1*b_1 + b_0
+sigma_i = 10
+y_1 = rnorm(n, xb, sigma_i)
+mean(xb)
+```
+
+```{r}
+  x_1 = rnorm(n, 10, 1)
+  b_1 = 1
+  b_0 = 0
+  xb = x_1*b_1 + b_0
+  sigma_i = 10
+  y_1 = rnorm(n, xb, sigma_i)
+  df_1_full = data.frame(y_1, x_1)
+
+  #Removing vals
+  five_percent = 1/(1+exp(-(df_1_full$x_1-6.5)))
+  mean(five_percent)
+  ten_percent = 1/(1+exp(-(df_1_full$x_1-7.45)))
+  mean(ten_percent)
+  twenty_percent = 1/(1+exp(-(df_1_full$x_1-8.35)))
+  mean(twenty_percent)
+  forty_percent = 1/(1+exp(-(df_1_full$x_1-9.51)))
+  mean(forty_percent)
+  sixty_percent = 1/(1+exp(-(df_1_full$x_1-10.49)))
+  mean(sixty_percent)
+  eighty_percent = 1/(1+exp(-(df_1_full$x_1-11.65)))
+  mean(eighty_percent)
+  ninety_percent = 1/(1+exp(-(df_1_full$x_1-12.52)))
+  mean(ninety_percent)
+```
+```{r}
+set.seed(1)
+  x_1 = rnorm(n, 10, 1)
+  b_1 = 1
+  b_0 = 0
+  xb = x_1*b_1 + b_0
+  sigma_i = 10
+  y_1 = rnorm(n, xb, sigma_i)
+  df_1_full = data.frame(y_1, x_1)
+
+  #Removing vals
+  five_percent = 1/(1+exp(-(df_1_full$x_1-6.5)))
+  mean(five_percent)
+  ten_percent = 1/(1+exp(-(df_1_full$x_1-7.45)))
+  mean(ten_percent)
+  twenty_percent = 1/(1+exp(-(df_1_full$x_1-8.35)))
+  mean(twenty_percent)
+  forty_percent = 1/(1+exp(-(df_1_full$x_1-9.51)))
+  mean(forty_percent)
+  sixty_percent = 1/(1+exp(-(df_1_full$x_1-10.49)))
+  mean(sixty_percent)
+  eighty_percent = 1/(1+exp(-(df_1_full$x_1-11.65)))
+  mean(eighty_percent)
+  ninety_percent = 1/(1+exp(-(df_1_full$x_1-12.52)))
+  mean(ninety_percent)
+  df_1_five = df_1_full
+  for (i in 1:length(five_percent)){
+    val = rbinom(1, 1, five_percent[i])
+    #print(val)
+    if(val == 0 ){
+      df_1_five$y_1[i] = NA
+    }
+  }
+
+  df_1_ten = df_1_full
+  for (i in 1:length(ten_percent)){
+    val = rbinom(1, 1, ten_percent[i])
+    #print(val)
+    if(val == 0 ){
+      df_1_ten$y_1[i] = NA
+    }
+  }
+
+  df_1_twenty = df_1_full
+  for (i in 1:length(twenty_percent)){
+    val = rbinom(1, 1, twenty_percent[i])
+    #print(val)
+    if(val == 0 ){
+      df_1_twenty$y_1[i] = NA
+    }
+  }
+
+  df_1_forty = df_1_full
+  for (i in 1:length(forty_percent)){
+    val = rbinom(1, 1, forty_percent[i])
+    #print(val)
+    if(val == 0 ){
+      df_1_forty$y_1[i] = NA
+    }
+  }
+
+  df_1_sixty = df_1_full 
+  for (i in 1:length(sixty_percent)){
+    val = rbinom(1, 1, sixty_percent[i])
+    #print(val)
+    if(val == 0 ){
+      df_1_sixty$y_1[i] = NA
+    }
+  }
+
+  df_1_eighty = df_1_full
+  for (i in 1:length(eighty_percent)){
+    val = rbinom(1, 1, eighty_percent[i])
+    #print(val)
+    if(val == 0 ){
+      df_1_eighty$y_1[i] = NA
+    }
+  }
+
+  df_1_ninety = df_1_full
+  for (i in 1:length(ninety_percent)){
+    val = rbinom(1, 1, ninety_percent[i])
+    #print(val)
+    if(val == 0 ){
+      df_1_ninety$y_1[i] = NA
+    }
+  }
+  colMeans(df_1_full)
+  colMeans(na.omit(df_1_five))
+  colMeans(na.omit(df_1_ten))
+  colMeans(na.omit(df_1_twenty))
+  colMeans(na.omit(df_1_forty))
+  colMeans(na.omit(df_1_sixty))
+  colMeans(na.omit(df_1_eighty))
+  colMeans(na.omit(df_1_ninety))
+```
+
+```{r}
+#Big run
+n = 700
+iters = 10
+imputations = 20
+#Empty files for appending results
+results_template =  data.frame("beta_0_mi" = double() , "beta_1_mi" = double(), "beta_1_mi_lower" = double(), "beta_1_mi_upper" = double(),  "beta_0_cca" = double(),  "beta_1_cca" = double(), "beta_1_cca_lower" =double(), "beta_1_cca_upper" =double(), "p-value_mi" = double(), "p-value_cca" = double())
+
+write.table(results_template, file='/home/marius/Downloads/Test/test_v2_5.csv', sep = ",", row.names=FALSE)
+write.table(results_template, file='/home/marius/Downloads/Test/test_v2_10.csv', sep = ",", row.names=FALSE)
+write.table(results_template, file='/home/marius/Downloads/Test/test_v2_20.csv', sep = ",", row.names=FALSE)
+write.table(results_template, file='/home/marius/Downloads/Test/test_v2_40.csv', sep = ",", row.names=FALSE)
+write.table(results_template, file='/home/marius/Downloads/Test/test_v2_60.csv', sep = ",", row.names=FALSE)
+write.table(results_template, file='/home/marius/Downloads/Test/test_v2_80.csv', sep = ",", row.names=FALSE)
+write.table(results_template, file='/home/marius/Downloads/Test/test_v2_90.csv', sep = ",", row.names=FALSE)
+
+for(i in 1:10){
+  print(i)
+  #Generating data according to Turid's recommended method
+  #One cont normal variables
+  set.seed(i)
+  x_1 = rnorm(n, 10, 1)
+  b_1 = 1
+  b_0 = 0
+  xb = x_1*b_1 + b_0
+  sigma_i = 10
+  y_1 = rnorm(n, xb, sigma_i)
+  df_1_full = data.frame(y_1, x_1)
+
+  #Removing vals
+  five_percent = 1/(1+exp(-(df_1_full$x_1-6.5)))
+  mean(five_percent)
+  ten_percent = 1/(1+exp(-(df_1_full$x_1-7.45)))
+  mean(ten_percent)
+  twenty_percent = 1/(1+exp(-(df_1_full$x_1-8.35)))
+  mean(twenty_percent)
+  forty_percent = 1/(1+exp(-(df_1_full$x_1-9.51)))
+  mean(forty_percent)
+  sixty_percent = 1/(1+exp(-(df_1_full$x_1-10.49)))
+  mean(sixty_percent)
+  eighty_percent = 1/(1+exp(-(df_1_full$x_1-11.65)))
+  mean(eighty_percent)
+  ninety_percent = 1/(1+exp(-(df_1_full$x_1-12.52)))
+  mean(ninety_percent)
+
+  removers = c(five_percent, ten_percent, twenty_percent, 
+               forty_percent, sixty_percent, eighty_percent, ninety_percent)
+
+  df_1_five = df_1_full
+  for (i in 1:length(five_percent)){
+    val = rbinom(1, 1, five_percent[i])
+    #print(val)
+    if(val == 0 ){
+      df_1_five$y_1[i] = NA
+    }
+  }
+
+  df_1_ten = df_1_full
+  for (i in 1:length(ten_percent)){
+    val = rbinom(1, 1, ten_percent[i])
+    #print(val)
+    if(val == 0 ){
+      df_1_ten$y_1[i] = NA
+    }
+  }
+
+  df_1_twenty = df_1_full
+  for (i in 1:length(twenty_percent)){
+    val = rbinom(1, 1, twenty_percent[i])
+    #print(val)
+    if(val == 0 ){
+      df_1_twenty$y_1[i] = NA
+    }
+  }
+
+  df_1_forty = df_1_full
+  for (i in 1:length(forty_percent)){
+    val = rbinom(1, 1, forty_percent[i])
+    #print(val)
+    if(val == 0 ){
+      df_1_forty$y_1[i] = NA
+    }
+  }
+
+  df_1_sixty = df_1_full 
+  for (i in 1:length(sixty_percent)){
+    val = rbinom(1, 1, sixty_percent[i])
+    #print(val)
+    if(val == 0 ){
+      df_1_sixty$y_1[i] = NA
+    }
+  }
+
+  df_1_eighty = df_1_full
+  for (i in 1:length(eighty_percent)){
+    val = rbinom(1, 1, eighty_percent[i])
+    #print(val)
+    if(val == 0 ){
+      df_1_eighty$y_1[i] = NA
+    }
+  }
+
+  df_1_ninety = df_1_full
+  for (i in 1:length(ninety_percent)){
+    val = rbinom(1, 1, ninety_percent[i])
+    #print(val)
+    if(val == 0 ){
+      df_1_ninety$y_1[i] = NA
+    }
+  }
+
+  #CCA
+  cca_1_5 = lm(y_1 ~x_1, data = df_1_five)
+  cca_1_10 = lm(y_1 ~ x_1, data = df_1_ten )
+  cca_1_20 = lm(y_1 ~ x_1, data = df_1_twenty)
+  cca_1_40 = lm(y_1 ~ x_1, data = df_1_forty)
+  cca_1_60 = lm(y_1 ~ x_1, data = df_1_sixty)
+  cca_1_80 = lm(y_1 ~ x_1, data = df_1_eighty)
+  cca_1_90 = lm(y_1 ~ x_1, data = df_1_ninety)
+
+  imp_1_5 = mice(df_1_five, m = imputations, maxit = iters, printFlag = FALSE)
+  imp_1_10 = mice(df_1_ten, m = imputations, maxit = iters, printFlag = FALSE)
+  imp_1_20 = mice(df_1_twenty, m = imputations, maxit = iters, printFlag = FALSE)
+  imp_1_40 = mice(df_1_forty, m = imputations, maxit = iters, printFlag = FALSE)
+  imp_1_60 = mice(df_1_sixty, m = imputations, maxit = iters, printFlag = FALSE)
+  imp_1_80 = mice(df_1_eighty, m = imputations, maxit = iters, printFlag = FALSE)
+  imp_1_90 = mice(df_1_ninety, m = imputations, maxit = iters, printFlag = FALSE)
+
+  fit_1_5 = with(imp_1_5, lm(y_1 ~x_1))
+  fit_1_10 = with(imp_1_10, lm(y_1 ~x_1))
+  fit_1_20 = with(imp_1_20, lm(y_1 ~x_1))
+  fit_1_40 = with(imp_1_40, lm(y_1 ~x_1))
+  fit_1_60 = with(imp_1_60, lm(y_1 ~x_1))
+  fit_1_80 = with(imp_1_80, lm(y_1 ~x_1))
+  fit_1_90 = with(imp_1_90, lm(y_1 ~x_1))
+
+  pooled_1_5 = pool(fit_1_5)
+  pooled_1_10 = pool(fit_1_10)
+  pooled_1_20 = pool(fit_1_20)
+  pooled_1_40 = pool(fit_1_40)
+  pooled_1_60 = pool(fit_1_60)
+  pooled_1_80 = pool(fit_1_80)
+  pooled_1_90 = pool(fit_1_90)
+
+  r_sq_1_5 = pool.r.squared(fit_1_5)
+  r_sq_1_10 = pool.r.squared(fit_1_10)
+  r_sq_1_20 = pool.r.squared(fit_1_20)
+  r_sq_1_40 = pool.r.squared(fit_1_40)
+  r_sq_1_60 = pool.r.squared(fit_1_60)
+  r_sq_1_80 = pool.r.squared(fit_1_80)
+  r_sq_1_90 = pool.r.squared(fit_1_90)
+
+  #Getting p-values
+  p_vals_mi = c(summary(pooled_1_5)[2,6], summary(pooled_1_10)[2,6],summary(pooled_1_20)[2,6],summary(pooled_1_40)[2,6],summary(pooled_1_60)[2,6],summary(pooled_1_80)[2,6],summary(pooled_1_90)[2,6])
+  p_vals_cca = c(summary(cca_1_5)$coefficients[2,4], summary(cca_1_10)$coefficients[2,4],summary(cca_1_20)$coefficients[2,4],summary(cca_1_40)$coefficients[2,4],summary(cca_1_60)$coefficients[2,4],summary(cca_1_80)$coefficients[2,4],summary(cca_1_90)$coefficients[2,4])
+
+  r_sq_cca = c(summary(cca_1_5)$r.squared, summary(cca_1_10)$r.squared , summary(cca_1_20)$r.squared,        summary(cca_1_40)$r.squared, summary(cca_1_60)$r.squared, summary(cca_1_80)$r.squared, summary(cca_1_90)$r.squared)
+  r_sq_total = c(r_sq_1_5[1], r_sq_1_10[1], r_sq_1_20[1], r_sq_1_40[1], r_sq_1_60[1], r_sq_1_80[1], r_sq_1_90[1])
+
+  beta_0_1_mi =c(summary(pooled_1_5)$estimate[1], summary(pooled_1_10)$estimate[1], summary(pooled_1_20)$estimate[1],   summary(pooled_1_40)$estimate[1], summary(pooled_1_60)$estimate[1], summary(pooled_1_80)$estimate[1],   summary(pooled_1_90)$estimate[1])
+
+  beta_1_1_mi =c(summary(pooled_1_5)$estimate[2], summary(pooled_1_10)$estimate[2], summary(pooled_1_20)$estimate[2],   summary(pooled_1_40)$estimate[2], summary(pooled_1_60)$estimate[2], summary(pooled_1_80)$estimate[2],   summary(pooled_1_90)$estimate[2])
+
+  beta_1_mi = cbind(beta_0_1_mi, beta_1_1_mi)
+
+  beta_1_mi_ci = rbind(beta_ci(summary(pooled_1_5,  conf.int=TRUE)), beta_ci(summary(pooled_1_10,  conf.int=TRUE)), beta_ci(summary(pooled_1_20,  conf.int=TRUE)), beta_ci(summary(pooled_1_40,  conf.int=TRUE)), beta_ci(summary(pooled_1_60,  conf.int=TRUE)), beta_ci(summary(pooled_1_80,  conf.int=TRUE)), beta_ci(summary(pooled_1_90,  conf.int=TRUE)))
+beta_1_mi_ci = data.frame("lower" = unlist(beta_1_mi_ci)[1:7], "upper" =unlist(beta_1_mi_ci)[8:14])
+
+  beta_0_1_cca_ci = rbind(confint(cca_1_5)[1,], confint(cca_1_10)[1,], confint(cca_1_20)[1,], confint(cca_1_40)[1,],   confint(cca_1_60)[1,], confint(cca_1_80)[1,], confint(cca_1_90)[1,])
+
+  beta_1_1_cca_ci = rbind(confint(cca_1_5)[2,], confint(cca_1_10)[2,], confint(cca_1_20)[2,], confint(cca_1_40)[2,],   confint(cca_1_60)[2,], confint(cca_1_80)[2,], confint(cca_1_90)[2,])
+  beta_both_1_cca = rbind(cca_1_5$coefficients, cca_1_10$coefficients, cca_1_20$coefficients, cca_1_40$coefficients,   cca_1_60$coefficients, cca_1_80$coefficients, cca_1_90$coefficients)
+  summary(cca_1_90)
+
+  df_1_beta = data.frame("beta_0_mi" = beta_0_1_mi , "beta_1_mi" = beta_1_1_mi, "beta_1_mi_lower" = beta_1_mi_ci$lower, "beta_1_mi_upper" = beta_1_mi_ci$upper,  "beta_0_cca" = beta_both_1_cca[,1], "beta_1_cca" = beta_both_1_cca[,2], "beta_1_cca_lower" =beta_1_1_cca_ci[,1],   "beta_1_cca_upper" =beta_1_1_cca_ci[,2], "p-value_mi" = p_vals_mi, "p-value_cca" = p_vals_cca)
+  #print(df_1_beta)
+
+  write.table(df_1_beta[1,], file = '/home/marius/Downloads/Test/test_v2_5.csv', sep = ",", append = TRUE, quote = FALSE,
+    col.names = FALSE, row.names = FALSE)
+  write.table(df_1_beta[2,], file = '/home/marius/Downloads/Test/test_v2_10.csv', sep = ",", append = TRUE, quote = FALSE,
+    col.names = FALSE, row.names = FALSE)
+  write.table(df_1_beta[3,], file = '/home/marius/Downloads/Test/test_v2_20.csv', sep = ",", append = TRUE, quote = FALSE,
+    col.names = FALSE, row.names = FALSE)
+  write.table(df_1_beta[4,], file = '/home/marius/Downloads/Test/test_v2_40.csv', sep = ",", append = TRUE, quote = FALSE,
+    col.names = FALSE, row.names = FALSE)
+  write.table(df_1_beta[5,], file = '/home/marius/Downloads/Test/test_v2_60.csv', sep = ",", append = TRUE, quote = FALSE,
+    col.names = FALSE, row.names = FALSE)
+  write.table(df_1_beta[6,], file = '/home/marius/Downloads/Test/test_v2_80.csv', sep = ",", append = TRUE, quote = FALSE,
+    col.names = FALSE, row.names = FALSE)
+  write.table(df_1_beta[7,], file = '/home/marius/Downloads/Test/test_v2_90.csv', sep = ",", append = TRUE, quote = FALSE,
+    col.names = FALSE, row.names = FALSE)
+
+}
+```
+
+```{r}
+p_results_5 = read.csv('/home/marius/Downloads/Test/v2_results_5.csv')
+p_results_5
+```
+
+```{r}
+#Loading experiment data
+results_5 = read.csv('/home/marius/Downloads/Test/test_v2_5.csv')
+results_10 = read.csv('/home/marius/Downloads/Test/test_v2_10.csv')
+results_20 = read.csv('/home/marius/Downloads/Test/test_v2_20.csv')
+results_40 = read.csv('/home/marius/Downloads/Test/test_v2_40.csv')
+results_60 = read.csv('/home/marius/Downloads/Test/test_v2_60.csv')
+results_80 = read.csv('/home/marius/Downloads/Test/test_v2_80.csv')
+results_90 = read.csv('/home/marius/Downloads/Test/test_v2_90.csv')
+
+#Checking amount of times we contain the true value
+tru_val = function(df){
+  mi = c()
+  cca = c()
+  for(i in 1:length(df[,1])){
+    upper_mi = df[i,4]
+    lower_mi = df[i,3]
+    t = FALSE
+    if(1<upper_mi&&lower_mi<1){
+      t = TRUE
+    }
+    if(t){
+      mi = c(mi, as.numeric(t))
+    }
+    else{
+      mi = c(mi, as.numeric(t))
+    }
+
+    upper_cca = df[i,8]
+    lower_cca = df[i,7]
+    t=FALSE
+    if(1<upper_cca&&lower_cca<1){
+      t = TRUE
+    }
+    if(t){
+      cca = c(cca, as.numeric(t))
+    }
+    else{
+      cca = c(cca, as.numeric(t))
+    }
+
+  }
+  return(list("mi" = mi,"cca" = cca))
+}
+
+tru_val_5 = tru_val(results_5)
+tru_val_10 = tru_val(results_10)
+tru_val_20 = tru_val(results_20)
+tru_val_40 = tru_val(results_40)
+tru_val_60 = tru_val(results_60)
+tru_val_80 = tru_val(results_80)
+tru_val_90 = tru_val(results_90)
+
+#Getting means
+a=data.frame(colMeans(results_5))
+r_5 = data.frame(t(a$colMeans.results_5.))
+colnames(r_5) = rownames(a)
+
+b=data.frame(colMeans(results_10))
+r_10 = data.frame(t(b$colMeans.results_10.))
+colnames(r_10) = rownames(b)
+
+c=data.frame(colMeans(results_20))
+r_20 = data.frame(t(c$colMeans.results_20.))
+colnames(r_20) = rownames(c)
+
+d=data.frame(colMeans(results_40))
+r_40 = data.frame(t(d$colMeans.results_40.))
+colnames(r_40) = rownames(d)
+
+e=data.frame(colMeans(results_60))
+r_60 = data.frame(t(e$colMeans.results_60.))
+colnames(r_60) = rownames(e)
+
+f=data.frame(colMeans(results_80))
+r_80 = data.frame(t(f$colMeans.results_80.))
+colnames(r_80) = rownames(f)
+
+g=data.frame(colMeans(results_90))
+r_90 = data.frame(t(g$colMeans.results_90.))
+colnames(r_90) = rownames(g)
+
+df_111 = rbind(r_5,r_10,r_20,r_40,r_60,r_80,r_90)
+df_111
+p<- ggplot(df_111, aes(x=c(5,10,20,40,60,80,90), y=beta_1_mi)) + 
+  geom_line() +
+  geom_point()+
+  geom_errorbar(aes(ymin=beta_1_mi_lower, ymax=beta_1_mi_upper), width=.2,
+                 position=position_dodge(0.05))
+t<- ggplot(df_111, aes(x=c(5,10,20,40,60,80,90), y=beta_1_cca)) + 
+  geom_line() +
+  geom_point()+
+  geom_errorbar(aes(ymin=beta_1_cca_lower, ymax=beta_1_cca_upper), width=.2,
+                 position=position_dodge(0.05))
+
+#Calculating bias
+mi_bias = c()
+cca_bias = c()
+for(i in 1:length(df_111[,1])){
+  bias_mi = 1-df_111[i,2] 
+  bias_cca = 1-df_111[i, 6]
+  mi_bias = c(mi_bias, bias_mi)
+  cca_bias = c(cca_bias, bias_cca)
+}
+cca_bias
+mi_bias
+t
+p
+
+#Calculating power
+
+#Calculating average size of CI
+ci_size = function(df){
+  mi = c()
+  cca = c()
+  for(i in 1:length(df[,1])){
+    upper_mi = df[i,4]
+    lower_mi = df[i,3]
+    mi = c(mi, upper_mi-lower_mi)
+
+    upper_cca = df[i,8]
+    lower_cca = df[i,7]
+
+    cca = c(cca, upper_cca-lower_cca)
+
+  }
+  return(list("mi" = mi,"cca" = cca))
+}
+
+ci_size_5 = ci_size(results_5)
+ci_size_5 = data.frame(ci_size_5)
+ci_size_10 = ci_size(results_10)
+ci_size_10 = data.frame(ci_size_10)
+ci_size_20 = ci_size(results_20)
+ci_size_20 = data.frame(ci_size_20)
+ci_size_40 = ci_size(results_40)
+ci_size_40 = data.frame(ci_size_40)
+ci_size_60 = ci_size(results_60)
+ci_size_60 = data.frame(ci_size_60)
+ci_size_80 = ci_size(results_80)
+ci_size_80 = data.frame(ci_size_80)
+ci_size_90 = ci_size(results_90)
+ci_size_90 = data.frame(ci_size_90)
+colMeans(data.frame(tru_val_5))
+colMeans(ci_size_40)
+```
+
+```{r}
+
+#One cont one binomial
+x_1 = rnorm(10, 1, n)
+x_2 = rbinom(n, 1, 0.5)
+
+b_1 = 1
+b_2 = 4
+
+xb = matrix(c(x_1*b_1, x_2*b_2), nrow = 2, ncol = 1)
+sigma_i = matrix(c(c(1,0), c(0,1)), nrow = 2, ncol = 2)
+
+y_2 = mvrnorm(n = n, xb, sigma_i, tol = 1e-14)
+
+```
+
+```{r}
+
+#One cont one binomial, 5 cont auxiliary
+x_1 = rnorm(10, 1, n)
+x_2 = rbinom(n, 1, 0.5)
+z_1 = rnorm(10,1,n)
+z_1 = rnorm(10,1,n)
+z_1 = rnorm(10,1,n)
+z_1 = rnorm(10,1,n)
+b_1 = 1
+b_2 = 0.8
+
+```
+
+```{r}
+
+#Simpler version with 3 variables correlated with Y
+sigma_test = matrix(c(c(1, 0.8, 0.4), c(0.8, 1, 0), c(0.4, 0, 1)), nrow = 3, ncol = 3)
+eigen(sigma_test)
+eigen(matrix(c(c(1, 0.8, 0.4, 0.6), c(0.8, 1, 0, 0), c(0.4, 0, 1, 0), c(0.6, 0, 0, 1)), nrow = 4, ncol = 4))
+mu_test = replicate(3, 0)
+simple_data = mvrnorm(n = n, mu_test, sigma_test, tol = 1e-14)
+df = data.frame(simple_data)
+
+#Playing around with simple model
+play_cov = matrix(c(c(10, sqrt(10)*0.8, sqrt(10)*0.4), c(sqrt(10)*0.8, 1, 0), c(sqrt(10)*0.4, 0, 1)), nrow = 3, ncol = 3)
+play_mu = c(180, 0, 0)
+play = mvrnorm(n = n, play_mu, play_cov, tol = 1e-14)
+df_p = data.frame(play)
+
+```
+
+```{r}
+#Removing data
+
+#Using the logit function to determine 
+test_logit = df$X2 + df$X3
+test_prob = 1/(1+exp(-(1.879+df$X2 + df$X3)))
+
+mean(test_prob)
+
+mar_list = c()
+for (item in test_prob){
+  mar_list = c(mar_list, c(rbinom(1, 100, item)))
+
+}
+mar_list
+mean(mar_list)
+test_prob[1]
+range(length(test_prob))
+#Removing values from Y
+for (i in 1:length(test_prob)){
+  val = rbinom(1, 1, test_prob[i])
+  #print(val)
+  if(val == 0 ){
+    df$X1[i] = NA
+  }
+}
+
+#Play
+play_prob = 1/(1+exp(-(-1.5+ df_p$X2 + df_p$X3)))
+mean(play_prob)
+df_p_full = df_p
+for (i in 1:length(play_prob)){
+  val = rbinom(1, 1, play_prob[i])
+  #print(val)
+  if(val == 0 ){
+    df_p$X1[i] = NA
+  }
+}
+df_p
+#Test
+#mylogit = glm(df$X1~ df$X2 + df$X3, data = df, family = "binomial")
+```
+
+```{r}
+#Multiple imputation
+imputed = mice(df)
+imputed$imp$X1
+
+fitm = with(imputed, lm(X1 ~ X2 + X3))
+summary(fitm)
+summary(pool(fitm))
+
+#Play
+imp_p = mice(df_p)
+imp_p$imp$X1
+
+fit_p = with(imp_p, lm(X1 ~ X2 + X3))
+summary(fit_p)
+summary(pool(fit_p))
+summary(pool(fit_p))
+pool.r.squared(fit_p)
+
+predict(pool(fit_p), data.frame(0,0))
+```
+
+```{r}
+#Inference
+#Test significance of regressions
+#Compare vs CCA
+#Test mu, sigma
+
+#Linear regression
+full_model = lm(X1 ~ X2 + X3, data = df_p_full)
+summary(full_model)
+full_model$coefficients[1] = 2
+summary(full_model)
+
+df_p_cca = na.omit(df_p)
+cca_model = lm(X1 ~ X2 + X3, data = df_p_cca)
+summary(cca_model)
+
+#Logistic regression
+```
+
+```{r}
+#Test for significance of logistic regression
+```
+
+```{r}
+#Whole set of regressions with varying levels of missing data
+
+#In the same graph compare CCA, MI, and True values.
+# We compare CI, Bias, ,
+
+#Wole set of regressions with varying levels of correlation
+
+#In the same graph compare CCA, MI, and True values
+
+#Whole set of regressions with categorical data as well.
+
+```
+
+```{r}
+#Testing regression with multivariate normal
+```

### 2020-06-11 final_presentation.tex
Finished

\documentclass{beamer}
\usepackage{tikz}
\usetikzlibrary{tikzmark,fit}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{xfrac}
\usepackage{float}
\usepackage{amssymb}
\usepackage[font={small}]{caption}
\usepackage{csquotes}
\usepackage{subfig}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{hyperref}
\captionsetup[figure]{labelfont=bf,textfont={it}}
\captionsetup[table]{position=above,labelfont=bf,textfont={it}}
\graphicspath{{figures/}}

%Information to be included in the title page:
\title{\textbf{Missing Data}}
\subtitle{\textbf{An Introduction}}
\author{Marius Johan Franco Dioli}
\date{June 2020}

\begin{document}

	\frame{\titlepage}

\begin{frame}
\frametitle{Table of Contents}
\tableofcontents
\end{frame}

\section{Defining missing data and missingness mechanisms}
\subsection{What is missing data?}
\begin{frame}
\frametitle{What is missing data?}
Missing data occurs when observations are missing in a data set. When we say an observation is missing, we mean that we have not observed it. The missing observation still has a value, but this value is hidden from us for some reason.
\end{frame}

\begin{frame}
\frametitle{Some basic notation}
\begin{itemize}
	\item<1-> Let the complete data matrix $Y=(y_{ij})$ be an $n \times k$ matrix, with the $i^{\text{th}}$ row $y_{i}=(y_{i1}, ..., y_{ik})$ where $y_{ij}$ is the value of variable $Y_{j}$ for observation $i$.
	\item<2->Let the missingness indicator matrix $M = (m_{ij})$ be an $n \times k$ matrix such that $m_{ij} = 1$ if $y_{ij}$ is missing and $m_{ij} = 0$ if $y_{ij}$ is observed.
	\item<3->Now, given $m_{i}$, we can partition $y_{i}$ into $y^{o}_{i}$ and $y^{m}_{i}$, corresponding to the components that are observed and missing respectively. 
\end{itemize}
\end{frame}

\subsection{Missingness mechanisms}
\begin{frame}
\frametitle{Missingness mechanisms}
The missingness mechanism is characterized by the conditional distribution of $m_{i}$ given $y_{i}$, i.e. $f_{M|Y}(m_{i}|y_{i}, \theta)$, where $\theta$ denotes unknown parameters. We distinguish between three different missingness mechanisms by considering how $M$ is related to $Y$ in order of decreasing strictness of assumptions: 
\end{frame}

\begin{frame}
\frametitle{Missing Completely At Random}
Missing Completely At Random (MCAR) occurs when, for all and any distinct values $y_{i}$ and $ y^{\ast}_{i}$ in the sample space of $Y$, $$f_{M|Y}(m_{i}|y_{i}, \theta) = f_{M|Y}(m_{i}|y^{\ast}_{i}, \phi), \quad \exists \theta, \phi$$ where $\theta$ and $\phi$ denote unknown parameters.
\end{frame}

\begin{frame}
\frametitle{Missing At Random}
Missing At Random (MAR) occurs when, for all i and any distinct values $y^{m}_{i}, y^{m\ast}_{i}$ in the sample space of $Y^{m}_{i}$, we have that the following equation holds: $$f_{M|Y}(m_{i}|y^{o}_{i}, y^{m}_{i}, \theta) = f_{M|Y}(m_{i}|y^{o}_{i}, y^{m\ast}_{i}, \phi), \qquad \exists \theta, \phi$$
\end{frame}

\begin{frame}
\frametitle{Missing Not At Random}
If $f_{M|Y}(m_{i}|y^{o}_{i}, y^{m}_{i}, \theta) = f_{M|Y}(m_{i}|y^{o}_{i}, y^{m\ast}_{i}, \phi)$ does not hold for some $i$ and some $y^{m}_{i}, y^{m\ast}_{i}$, then the missingness mechanism is called Missing Not At Random (MNAR).
\end{frame}

\subsection{Illustrative example}
\begin{frame}
\frametitle{}
\begin{table}[]
	\begin{tabular}{ccccc}
		\multicolumn{1}{l}{2018} & \multicolumn{1}{l}{2019 (Complete)}& \multicolumn{1}{l}{2019 (MCAR)} & \multicolumn{1}{l}{2019 (MAR)} & \multicolumn{1}{l}{2019 (MNAR)} \\ \hline
		8.9                      & 7.4                                 & 7.4                             & 7.4                            & NA                              \\
		9.4                      & 7.3                                 & 7.3                             & 7.3                            & NA                              \\
		8.4                      & 7.9                                 & NA                              & 7.9                            & NA                              \\
		5.3                      & 6                                   & NA                              & NA                             & NA                              \\
		11.1                     & 7.4                                 & 7.4                             & 7.4                            & NA                              \\
		8.4                      & 10.7                                & NA                              & 10.7                           & 10.7                            \\
		6.4                      & 8.8                                 & NA                              & NA                             & 8.8                             \\
		5.6                      & 3.5                                 & NA                              & NA                             & NA                              \\
		5.6                      & 7.1                                 & 7.1                             & NA                             & NA                              \\
		8.2                      & 6.9                                 & 6.9                             & 6.9                            & NA                              
	\end{tabular}

\end{table}
\end{frame}

\begin{frame}
\frametitle{}
\begin{table}[]
	\begin{tabular}{ccccc}
		\multicolumn{1}{l}{2018} & \multicolumn{1}{l}{2019 (Complete)}& \multicolumn{1}{l}{2019 (MCAR)} & \multicolumn{1}{l}{2019 (MAR)} & \multicolumn{1}{l}{2019 (MNAR)} \\ \hline
		8.9                      & 7.4                                 & 7.4                             & 7.4                            & NA                              \\
		9.4                      & 7.3                                 & 7.3                             & 7.3                            & NA                              \\
		8.4                      & 7.9                                 & NA                              & 7.9                            & NA                              \\
		5.3                      & 6                                   & NA                              & NA                             & NA                              \\
		11.1                     & 7.4                                 & 7.4                             & 7.4                            & NA                              \\
		8.4                      & 10.7                                & NA                              & 10.7                           & 10.7                            \\
		6.4                      & 8.8                                 & NA                              & NA                             & 8.8                             \\
		5.6                      & 3.5                                 & NA                              & NA                             & NA                              \\
		5.6                      & 7.1                                 & 7.1                             & NA                             & NA                              \\
		8.2                      & 6.9                                 & 6.9                             & 6.9                            & NA                              
	\end{tabular}

\end{table}
\end{frame}

\begin{frame}
\frametitle{}
\begin{table}[]
	\begin{tabular}{ccccc}
		\multicolumn{1}{l}{2018} & \multicolumn{1}{l}{2019 (Complete)}& \multicolumn{1}{l}{2019 (MCAR)} & \multicolumn{1}{l}{2019 (MAR)} & \multicolumn{1}{l}{2019 (MNAR)} \\ \hline
		8.9                      & 7.4                                 & 7.4                             & 7.4                            & NA                              \\
		9.4                      & 7.3                                 & 7.3                             & 7.3                            & NA                              \\
		8.4                      & 7.9                                 & NA                              & 7.9                            & NA                              \\
		5.3                      & 6                                   & NA                              & NA                             & NA                              \\
		11.1                     & 7.4                                 & 7.4                             & 7.4                            & NA                              \\
		8.4                      & 10.7                                & NA                              & 10.7                           & 10.7                            \\
		6.4                      & 8.8                                 & NA                              & NA                             & 8.8                             \\
		5.6                      & 3.5                                 & NA                              & NA                             & NA                              \\
		5.6                      & 7.1                                 & 7.1                             & NA                             & NA                              \\
		8.2                      & 6.9                                 & 6.9                             & 6.9                            & NA                              
	\end{tabular}

\end{table}
\end{frame}

\begin{frame}
\frametitle{}
\begin{table}[]
	\begin{tabular}{ccccc}
		\multicolumn{1}{l}{2018} & \multicolumn{1}{l}{2019 (Complete)}& \multicolumn{1}{l}{2019 (MCAR)} & \multicolumn{1}{l}{2019 (MAR)} & \multicolumn{1}{l}{2019 (MNAR)} \\ \hline
		8.9                      & 7.4                                 & 7.4                             & 7.4                            & NA                              \\
		9.4                      & 7.3                                 & 7.3                             & 7.3                            & NA                              \\
		8.4                      & 7.9                                 & NA                              & 7.9                            & NA                              \\
		5.3                      & 6                                   & NA                              & NA                             & NA                              \\
		11.1                     & 7.4                                 & 7.4                             & 7.4                            & NA                              \\
		8.4                      & 10.7                                & NA                              & 10.7                           & 10.7                            \\
		6.4                      & 8.8                                 & NA                              & NA                             & 8.8                             \\
		5.6                      & 3.5                                 & NA                              & NA                             & NA                              \\
		5.6                      & 7.1                                 & 7.1                             & NA                             & NA                              \\
		8.2                      & 6.9                                 & 6.9                             & 6.9                            & NA                              
	\end{tabular}

\end{table}
\end{frame}

\section{Imputation}
\begin{frame}
\frametitle{What is imputation?}
Imputation refers to the substituting, or \enquote{filling in}, of missing values with plausible estimates. This is done with goal of improving the subsequent analysis. Most analysis techniques take as input complete data, and so imputation complements this by creating a complete (though not \textit{the} complete) data set.
\end{frame}

\subsection{Congeniality}
\begin{frame}
\frametitle{Congeniality}
When the analysis model differs substantially from the imputation model we say that they are uncongenial. Such differences might cause the subsequent analysis to be biased, or in the worst case, invalid. 
\end{frame}

\section{Ad hoc methods}
\begin{frame}
\frametitle{Ad hoc methods}
Although ad hoc is pejorative in certain circumstances, it is used here to indicate that the methods below are motivated by convenience rather than any methodological concern. Below is a simplified version of the earlier example.

\begin{table}[]
	\begin{tabular}{lllllll}
		& T\_0 & T\_1 & T\_2 & T\_3 & T\_4 & T\_5 \\
		Person A & 8.1  & 8.6  & 7    & 9    & 8.2  & 8.0  \\
		Person B & 6    & 5.2  & NA   & NA   & NA   & NA 
	\end{tabular}
\end{table}

\end{frame}

\subsection{Complete Case Analysis}
\begin{frame}
\frametitle{Complete Case Analysis}
Complete Case Analysis (CCA) involves restricting analysis to entries with complete data. CCA holds an intuitive appeal, as it is both exceedingly simple to implement, and with the only seeming downside being a loss in accuracy. However, if the missingness mechanism is not MCAR, then CCA will lead to distortions in parameter estimates and their variances\cite{Schafer2002}.

\begin{table}[]
	\begin{tabular}{lllllll}
		& T\_0 & T\_1 & T\_2 & T\_3 & T\_4 & T\_5 \\
		Person A & 8.1  & 8.6  & 7    & 9    & 8.2  & 8.0 
	\end{tabular}
\end{table}

\end{frame}

\subsection{Mean imputation}
\begin{frame}
\frametitle{Mean imputation}
Mean imputation involves substituting the missing values with the mean of the observed values for that variable. As stated in Schafer and Graham \cite{Schafer2002}, this has the effect of preserving the mean of the distribution one wishes to draw inference about, but distorts other parameters such as variance and covariance.

\begin{table}[]
	\begin{tabular}{lllllll}
		& T\_0 & T\_1 & T\_2 & T\_3 & T\_4 & T\_5 \\
		Person A & 8.1  & 8.6  & 7    & 9    & 8.2  & 8.0  \\
		Person B & 6    & 5.2  & 5.6   & 5.6   & 5.6   & 5.6  
	\end{tabular}
\end{table}

\end{frame}

\subsection{Last Observation Carried Forward}
\begin{frame}
\frametitle{Last Observation Carried Forward}
Last Observation Carried Forward (LOCF) involves imputing the missing values for an individual with the last observed value for that individual. By replicating already existing values, LOCF has a dampening impact on the variance estimate.

\begin{table}[]
	\begin{tabular}{lllllll}
		& T\_0 & T\_1 & T\_2 & T\_3 & T\_4 & T\_5 \\
		Person A & 8.1  & 8.6  & 7    & 9    & 8.2  & 8.0  \\
		Person B & 6    & 5.2  & 10   & 10   & 10   & 10  
	\end{tabular}
\end{table}

\end{frame}

\subsection{Regression Imputation}
\begin{frame}
\frametitle{Regression Imputation}
Regression imputation functions by regressing the variable(s) with missing entries on the fully observed variables, or subsets that are fully observed. This is the best of the ad hoc methods, as it gives us much more plausible imputations that vary more like the true data \cite[P. 37]{Molenberghs2015}. However, even in this simple case it yields imputations that tend to be less variable than the observed data. Therefore, using these imputations in inference will most often lead to underestimating standard errors and p-values. If there are non-linear relationships in the data, this method's performance breaks down.
\end{frame}

\section{Multiple Imputation}
\begin{frame}
\frametitle{Multiple Imputation}
Multiple Imputation (MI) is the process of creating multiple distinct imputed data sets, running our analysis, and then combining the results. It functions in the following way:
\begin{enumerate}
	\item<1-> $M$ imputed datasets are generated through repeated random draws from the predictive distribution of the missing values, $f(Y^{m}|Y^{o}, M)$, under a particular model for missingness.
	\item<2-> The intended analysis is performed separately on all $M$ datasets
	\item<3-> The results of the analyses are combined using Rubin's or Meng's rules to create a single MI result.
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{Combining rules}
At step 2 of MI, we have $M$ sets of parameter estimates and their variances: $\hat{\theta}_{i},\hat{Var_{i}}, \quad i=0,...,M$. The parameter estimate is simply the average $$\bar{\theta} = \frac{1}{M}\sum_{i=0}^{M}\hat{\theta}_{i}$$ and the variance associated with $\bar{\theta}$ is given by

\begin{equation}
T = \overline{Var}+(1+\frac{1}{M}B)\tag{*}
\end{equation}

\end{frame}

\begin{frame}
$$\overline{Var} = \frac{1}{M}\sum_{i=0}^{M}\hat{Var}_{i}$$ measures the within imputation variability and $$B = \frac{1}{M-1}\sum_{i=0}^{M}(\hat{\theta_{i}}-\bar{\theta})(\hat{\theta_{i}}-\bar{\theta})^{T}$$ measures the between imputation variability. The adjustment $(1+\frac{1}{M}B)$ in equation $(*)$ is due to the finite number of imputations. If we let $M\rightarrow\infty$ then we achieve asymptotic equality to the true variance estimate.
\end{frame}

\begin{frame}
\frametitle{Strengths and weaknesses of MI}
Strengths:
\begin{enumerate}
	\item<1-> By modeling the missing variables and combining multiple imputed datasets, MI achieves better variance estimates.
	\item<2-> Can model different missingness mechanisms.
	\item<3-> Can capture complex relationships between the observed and the missing data.
\end{enumerate}

Weaknesses:
\begin{enumerate}
	\item<1-> Computationally expensive.
	\item<2-> Specifying the imputation model can be difficult, for example if the there is a mix of variable types and a hierarchical structure in the data.
	\item<3-> Improper specification of the imputation model can cause uncongeniality. 
\end{enumerate}
\end{frame}

\section{Experiments}
\subsection{Experimental setup}
\begin{frame}
\frametitle{Experimental setup}
We looked at a simple linear regression with a single covariate defined by the model $$Y_{i} = \beta_{0} + \beta_{1}x_{i} + \epsilon, \qquad \epsilon \sim N(0,1) \qquad i\in(0,...,n)$$. Missingness was be simulated in both $X$ and $Y$. Data was generated by drawing from a bivariate normal distribution $$[Y X] \sim N(\begin{bmatrix}
0 \\
0 \\
\end{bmatrix},\begin{bmatrix}
1 & 0.8 \\
0.8 & 1 \\
\end{bmatrix})$$ The true regression coefficient is then $\beta_{1} = 0.8$. In our experiment we drew 1000 data points from this distribution. MCAR was simulated by removing the first p values in the data set. MAR and MNAR were simulated using the logit function. The number of imputations $M$ was set to 20.
\end{frame}

\begin{frame}
\frametitle{Experimental setup}
We will look at the following metrics to evaluate the performance of MI:

\begin{enumerate}
	\item<1-> Size of the confidence interval of $\beta_{1}$
	\item<2-> Coverage of the confidence interval
	\item<3-> Bias
\end{enumerate}

We will compare the performance of MI to an equivalent analysis performed with CCA for reference. Our experiment was then conducted in the following manner:

\begin{enumerate}
	\item<4-> Draw 1000 observations from the distributions of $[Y X]$
	\item<5-> Simulate 5, 10, 20, 40, 60, 80, and 90 percent missingness under different missingness mechanisms.
	\item<6-> Create linear models for each of these 7 datasets based on CCA and MI
	\item<7-> Repeat 1000 times
	\item<8-> Evaluate the results based on the above metrics.
\end{enumerate}
\end{frame}

\subsection{Results}
\begin{frame}
\frametitle{Results}
\begin{figure}[H]
	\centering
	\subfloat[$\beta_{1}$ estimate with CCA]{\includegraphics[width=0.3\textwidth]{mcar_estimate_cca.png}}
	\quad
	\subfloat[$\beta_{1}$ estimate with MI]{\includegraphics[width=0.3\textwidth]{mcar_estimate_mi.png}}
	\\
	\subfloat[Coverage]{\includegraphics[width=0.2\textwidth]{mcar_coverage.png}}
	\quad
	\subfloat[Average CI size]{\includegraphics[width=0.2\textwidth]{mcar_ci.png}}
	\quad
	\subfloat[Average bias]{\includegraphics[width=0.2\textwidth]{mcar_bias.png}}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{MAR in X}
\begin{figure}[H]
	\centering
	\subfloat[$\beta_{1}$ estimate with CCA]{\includegraphics[width=0.3\textwidth]{mar_reg_x_estimate_cca.png}}
	\quad
	\subfloat[$\beta_{1}$ estimate with MI]{\includegraphics[width=0.3\textwidth]{mar_reg_x_estimate_mi.png}}
	\\
	\subfloat[Coverage]{\includegraphics[width=0.2\textwidth]{mar_reg_x_coverage.png}}
	\quad
	\subfloat[Average CI size]{\includegraphics[width=0.2\textwidth]{mar_reg_x_ci.png}}
	\quad
	\subfloat[Average bias]{\includegraphics[width=0.2\textwidth]{mar_reg_x_bias.png}}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{MAR in Y}
\begin{figure}[H]
	\centering
	\subfloat[$\beta_{1}$ estimate with CCA]{\includegraphics[width=0.3\textwidth]{final_estimate_cca.png}}
	\quad
	\subfloat[$\beta_{1}$ estimate with MI]{\includegraphics[width=0.3\textwidth]{final_estimate_mi.png}}
	\\
	\subfloat[Coverage]{\includegraphics[width=0.2\textwidth]{final_coverage.png}}
	\quad
	\subfloat[Average CI size]{\includegraphics[width=0.2\textwidth]{final_ci.png}}
	\quad
	\subfloat[Average bias]{\includegraphics[width=0.2\textwidth]{final_bias.png}}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{MNAR}
\begin{figure}[H]
	\centering
	\subfloat[$\beta_{1}$ estimate with CCA]{\includegraphics[width=0.3\textwidth]{mnar_final_estimate_cca.png}}
	\quad
	\subfloat[$\beta_{1}$ estimate with MI]{\includegraphics[width=0.3\textwidth]{mnar_final_estimate_mi.png}}
	\\
	\subfloat[Coverage]{\includegraphics[width=0.2\textwidth]{mnar_final_coverage.png}}
	\quad
	\subfloat[Average CI size]{\includegraphics[width=0.2\textwidth]{mnar_final_ci.png}}
	\quad
	\subfloat[Average bias]{\includegraphics[width=0.2\textwidth]{mnar_final_bias.png}}
\end{figure}
\end{frame}

\section{Conclusion and further work}
\begin{frame}
\frametitle{Conclusion and further work}
We conclude that:
\begin{enumerate}
	\item<1-> Missing data is an important consideration in data analysis.
	\item<2-> Imputation can in certain circumstances help with missing data.
	\item<3-> Theory and literature disagree with experimental results. Further study will be needed to verify these results.
\end{enumerate}
\end{frame}

\begin{frame}
\frametitle{Conclusion and further work}
We conclude that:
\begin{enumerate}
	\item Missing data is an important consideration in data analysis.
	\item Imputation can in certain circumstances help with missing data.
	\item Theory and literature disagree with experimental results. Further study will be needed to verify these results.
\end{enumerate}
Possible future investigations could be made into the effect of MI on prediction and classification, especially in the context of machine learning models. Another interesting avenue of exploration is replacing the conditional distributions with other modeling tools such as trees or neural networks. Perhaps the messy nature of much of modern data will make the underlying assumptions of conditional distributions untenable.
\end{frame}

\bibliography{final_5_bib}	
\bibliographystyle{apalike}
\end{document}
